---
title: "Wastewater Growth of Ulva and Hypnea"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Data collection was divided into Ulva and Hypnea
### Dictionary of terms: 
#### plant_id = plants were divided and placed into two or more treatments/salinities,
#### treatment = nutrient dose, which includes nitrate and phosphate,
#### growth_d9 = calculated growth from final weight-initial weight divided by initial,
#### run = period during which each experiment was conducted. There are six runs,
#### lunar_phase = for Ulva lactuca, lunar phase important for tissue sloughing


# Load packages for mixed model effects analysis
```{r}
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
library(gt)
library(purrr)
library(stringr)
library(tidyr)
library(piecewiseSEM)
library(easystats)
library(magrittr)
```

# Open weight dataset and make columns for growth rate from initial and final weights
```{r}
waste_p2_growth <- read.csv("../../data/wastewater/input/growth_2025/waste_p2_growth.csv")
```

# Assigns treatment as characters from integers then to factors
```{r}
waste_p2_growth$treatment <- as.factor(as.character(waste_p2_growth$treatment))
waste_p2_growth$salinity <- as.factor(as.character(waste_p2_growth$salinity))
```


# Create data subsets for the analysis and plots
# Use Day 9 for final analysis
```{r}
ulva_g <- subset(waste_p2_growth, species == "u")
ulva_g$treatment_graph[ulva_g$treatment == 1] <- "1) 80 umol"
ulva_g$treatment_graph[ulva_g$treatment == 3] <- "2) 245 umol" 
ulva_g$treatment_graph[ulva_g$treatment == 5] <- "3) 748umol"
ulva_g$treatment_graph[ulva_g$treatment == 7] <- "4) 2287 umol"
```

```{r}
hypnea_g <- subset(waste_p2_growth, species == "h")
hypnea_g$treatment_graph[hypnea_g$treatment == 1] <- "1) 80 umol"
hypnea_g$treatment_graph[hypnea_g$treatment == 3] <- "2) 245 umol" 
hypnea_g$treatment_graph[hypnea_g$treatment == 5] <- "3) 748 umol"
hypnea_g$treatment_graph[hypnea_g$treatment == 7] <- "4) 2287 umol"
```

# Make histograms
```{r}
ulva_g %>% ggplot(aes(growth_d9)) +
  geom_histogram(binwidth=5, fill = "darkolivegreen3", color = "black", linewidth = 0.25, alpha = 0.85) +
  theme_bw()
```

# Run the model (removed lunar_phase since this had no effect and caused issues of Singularity)
```{r}
ulva_growth_model <- lmer(formula = growth_d9 ~ 0 + salinity + treatment +
                              (1 | plant_id) + (1 | run), data = ulva_g, REML = TRUE)
```

#Check basic performance
```{r}
isSingular(ulva_growth_model)
hist(resid(ulva_growth_model))
plot(resid(ulva_growth_model) ~ fitted(ulva_growth_model))
qqnorm(resid(ulva_growth_model))
qqline(resid(ulva_growth_model))
```

# Check the performance of the model for dataset: ulva
```{r}
performance::check_model(ulva_growth_model)
r.squaredGLMM(ulva_growth_model)
summary(ulva_growth_model)
emms_t <- emmeans(ulva_growth_model, ~treatment)
summary (emms_t)
emms_s <- emmeans(ulva_growth_model, ~salinity)
summary(emms_s)
```

# View random effects levels
```{r}
tab_model(ulva_growth_model, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(ulva_growth_model))
```

# Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
ulva_treatment_null <- lmer(formula = growth_d9 ~ salinity + 
                              (1 | plant_id) + (1 | run) + (1 | lunar_phase), data = ulva_g, REML = FALSE)
ulva_with_treatment <- lmer(formula = growth_d9 ~ treatment + salinity +
                        (1 | plant_id) + (1 | run) + (1 | lunar_phase), data = ulva_g, REML = FALSE)
anova(ulva_treatment_null, ulva_with_treatment)
ulva_salinity_null <- lmer(formula = growth_d9 ~ treatment + (1 | plant_id) + (1 | run) + (1 | lunar_phase), data = ulva_g, REML = FALSE)
ulva_with_salinity <- lmer(formula = growth_d9 ~ treatment + salinity + (1 | plant_id) + (1 | run) + (1 | lunar_phase), data = ulva_g, REML = FALSE)
anova(ulva_salinity_null, ulva_with_salinity)
```

# Plot ulva growth with treatment on x-axis
```{r}
ulva_p2_growth_plot <- ulva_g %>% 
  ggplot(aes(treatment_graph, growth_d9)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 2, aes(color = salinity), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "9-Day Growth (%)", title= "A - Growth", subtitle = "Ulva lactuca") + 
  scale_x_discrete(labels = c("80 μmol N", "245 μmol N", 
                              "748 μmol N", "2287 μmol N")) + 
  ylim(-30, 370) + stat_mean() + 
  scale_color_manual(values = c("18" = "chartreuse4","22" = "darkolivegreen2")) +
  geom_hline(yintercept=125, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1), legend.background = element_rect(fill = "white", color = "black"),
        plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
ulva_p2_growth_plot
```

# Plot for growth with salinity on x-axis
```{r}
ulva_p2_growth_plot2 <- ulva_g %>% 
  ggplot(aes(salinity, growth_d9)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 2, aes(color = treatment), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "9-Day Growth (%)", title= "B - Growth", subtitle = "Ulva lactuca") + 
  scale_x_discrete(labels = c("18 ppt", "22 ppt")) + 
  ylim(-30, 370) + stat_mean() + 
  #scale_color_manual(values = c("deeppink", "maroon4", "darkred", "firebrick1")) +
  geom_hline(yintercept=64, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1), legend.background = element_rect(fill = "white", color = "black"),
        plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
ulva_p2_growth_plot2
```

# Summarize the means
```{r}
waste_p2_growth %>% group_by(species, treatment) %>% summarise_at(vars(growth_d9), list(mean = mean))
waste_p2_growth %>% group_by(species) %>% summarise_at(vars(growth_d9), list(mean = mean))
```

## Hypnea

# Make a histogram of the growth rate data
```{r}
hypnea_g %>% ggplot(aes(growth_d9)) +
  geom_histogram(binwidth=5, fill = "deeppink4", color = "black", linewidth = 0.25, alpha = 0.85) +
  theme_bw()
```



# Run growth model for Hypnea
```{r}
growth_model_hypnea <- lmer(formula = growth_d9 ~ salinity + treatment +
                             (1 | plant_id) + (1 | run), data = hypnea_g, REML = TRUE)
```

# Check basic performance
```{r}
isSingular(growth_model_hypnea)
hist(resid(growth_model_hypnea))
plot(resid(growth_model_hypnea) ~ fitted(growth_model_hypnea))
qqnorm(resid(growth_model_hypnea))
qqline(resid(growth_model_hypnea))
```

# Check the performance of the model for dataset: hypnea
```{r}
performance::check_model(growth_model_hypnea)
r.squaredGLMM(growth_model_hypnea)
summary(growth_model_hypnea)
tab_model(growth_model_hypnea, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(growth_model_hypnea))
```

# Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
hypnea_nitrate_null <- lmer(formula = growth_d9 ~ salinity + 
                             (1 | plant_id) + (1 | run), data = hypnea_g, REML = FALSE)
hypnea_model2 <- lmer(formula = growth_d9 ~ treatment + salinity +
                       (1 | plant_id) + (1 | run), data = hypnea_g, REML = FALSE)
anova(hypnea_nitrate_null, hypnea_model2)
hypnea_salinity_null <- lmer(formula = growth_d9 ~ treatment + (1 | plant_id) + (1 | run), data = hypnea_g, REML = FALSE)
hypnea_model3 <- lmer(formula = growth_d9 ~ treatment + salinity + (1 | plant_id) + (1 | run), data = hypnea_g, REML = FALSE)
anova(hypnea_salinity_null, hypnea_model3)
```

# Plot hypnea growth
```{r}
hypnea_p2_growth_plot <- hypnea_g %>% 
  ggplot(aes(treatment_graph, growth_d9)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 2, aes(color = salinity), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "9-Day Growth (%)", title= "B - Growth", subtitle = "Hypnea musciformis") + 
  scale_x_discrete(labels = c("80 μmol N", "245 μmol N",
                              "748 μmol N", "2287 μmol N")) + 
  ylim(0, 150) + stat_mean() + 
  scale_color_manual(values = c("deeppink", "maroon4")) +
  geom_hline(yintercept=64, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1), legend.background = element_rect(fill = "white", color = "black"),
        plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
hypnea_p2_growth_plot
```

# Plot for growth with salinity on x-axis
```{r}
hypnea_p2_growth_plot2 <- hypnea_g %>% 
  ggplot(aes(salinity, growth_d9)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 2, aes(color = treatment), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "9-Day Growth (%)", title= "B - Growth", subtitle = "Hypnea musciformis") + 
  scale_x_discrete(labels = c("18 ppt", "22 ppt")) + 
  ylim(-0, 160) + stat_mean() + 
  #scale_color_manual(values = c("deeppink", "maroon4", "darkred", "firebrick1")) +
  geom_hline(yintercept=64, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1), legend.background = element_rect(fill = "white", color = "black"),
        plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
hypnea_p2_growth_plot2
```





