---
title: "Spectral Reflectance — Ulva lactuca & Hypnea musciformis"
subtitle: "Wastewater SGD 2025 Experiment"
author: "Angela Richards Dona"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 9,
  fig.height = 5
)
```

```{r libraries}
library(tidyverse)
library(scales)
library(grid)
```

```{r load-data}
spec_norm <- read_csv(
  "/Users/angela/src/Photosynthesis/data/wastewater/transformed/spec_norm_long.csv",
  show_col_types = FALSE
) %>%
  mutate(
    nitrate  = factor(nitrate,  levels = sort(unique(nitrate))),
    salinity = factor(salinity, levels = sort(unique(salinity))),
    species  = as.character(species)
  )

spec_u <- filter(spec_norm, species == "u")
spec_h <- filter(spec_norm, species == "h")
```

```{r peak-annotations}
# Species-specific pigment peak wavelengths (observed in spectra)
peaks_ulva <- tibble(
  wavelength = c(435, 500, 572, 650, 675),
  label      = c("Chl a/b + Car", "Carotenoid", "Green trough", "Chl b", "Chl a")
)

peaks_hypnea <- tibble(
  wavelength = c(435, 500, 572, 625, 675),
  label      = c("Chl a/b + Car", "B-PE", "R-PE", "PC", "Chl a")
)
```

---

## Overview

Spectral absorptance was measured on Day 9 of the wastewater SGD 2025 experiment for *Ulva lactuca* (green alga) and *Hypnea musciformis* (red alga) across four nitrate concentrations (80, 245, 748, 2287 µmol) and two salinities (18 and 22 ppt).

Raw reflectance was smoothed (Savitzky-Golay, window = 11, polynomial order = 2), converted to absorptance (100 − reflectance), normalized by SNV (standard normal variate), and baseline-corrected to zero at 749 nm.

**Note on Ulva spectra:** The spectrometer range ends at ~750.7 nm. For *Ulva*, the red edge reflectance peaks at ~688–691 nm and then *declines* to 749 nm — the true NIR plateau is not captured. As a result, the 749 nm baseline correction pushes the 680–700 nm region negative in the processed spectra. The dominant absorptance peak for *Ulva* in these specimens is at ~650 nm (Chl b), consistent with the observed reflectance minimum in the raw data. This is a real spectral characteristic of these specimens, not a processing artifact.

Key absorptance peaks annotated (observed wavelengths):

| Species | Wavelength | Pigment / Feature |
|---------|-----------|---------|
| Both | 435 nm | Chl a/b + carotenoids (Soret band) |
| *Ulva* | 500 nm | Carotenoid / accessory pigment absorption |
| *Hypnea* | 500 nm | B-phycoerythrin Soret peak (observed) |
| *Ulva* | 572 nm | Green reflectance trough |
| *Hypnea* | 572 nm | R-phycoerythrin absorption peak (observed; shifted from textbook 560 nm) |
| *Hypnea* | 625 nm | Phycocyanin (observed; shifted from textbook 620 nm) |
| *Ulva* | 650 nm | Chl b red absorption peak (dominant red peak in these specimens) |
| Both | 675 nm | Chl a red absorption maximum |

---

## 1. Individual Spectra

### *Ulva lactuca*

```{r ulva-individual, fig.height=5}
ggplot(spec_u, aes(x = wavelength, y = absorptance_zero, color = id)) +
  geom_vline(data = peaks_ulva, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey60", linewidth = 0.35, inherit.aes = FALSE) +
  geom_line(linewidth = 0.3, alpha = 0.8) +
  facet_wrap(salinity ~ nitrate, nrow = 2,
             labeller = labeller(
               salinity = function(x) paste0(x, " ppt"),
               nitrate  = function(x) paste0(x, " µmol")
             )) +
  xlim(400, 750) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(
    x       = "Wavelength (nm)",
    y       = "Zero-baseline absorptance (SNV)",
    title   = expression(italic("Ulva lactuca") ~ "— Individual Spectra by Treatment"),
    caption = "Dashed lines: 435 nm | 500 nm Carotenoid | 572 nm Green trough | 650 nm Chl b | 675 nm Chl a"
  )
```

### *Hypnea musciformis*

```{r hypnea-individual, fig.height=5}
ggplot(spec_h, aes(x = wavelength, y = absorptance_zero, color = id)) +
  geom_vline(data = peaks_hypnea, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey60", linewidth = 0.35, inherit.aes = FALSE) +
  geom_line(linewidth = 0.3, alpha = 0.8) +
  facet_wrap(salinity ~ nitrate, nrow = 2,
             labeller = labeller(
               salinity = function(x) paste0(x, " ppt"),
               nitrate  = function(x) paste0(x, " µmol")
             )) +
  xlim(400, 750) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(
    x       = "Wavelength (nm)",
    y       = "Zero-baseline absorptance (SNV)",
    title   = expression(italic("Hypnea musciformis") ~ "— Individual Spectra by Treatment"),
    caption = "Dashed lines: 435 nm | 500 nm B-PE | 572 nm R-PE | 625 nm PC | 675 nm Chl a"
  )
```

---

## 2. Mean Spectra by Nitrate Treatment

Mean ± SE absorptance across all replicates, grouped by nitrate concentration. Salinity is pooled to highlight the nitrate signal.

```{r mean-spectra-nitrate}
spec_summ_nitrate <- spec_norm %>%
  filter(wavelength >= 400, wavelength <= 750) %>%
  group_by(species, nitrate, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    se_abs   = sd(absorptance_zero,   na.rm = TRUE) / sqrt(n()),
    .groups  = "drop"
  ) %>%
  mutate(
    species_label = recode(species, u = "Ulva lactuca", h = "Hypnea musciformis")
  )
```

### *Ulva lactuca*

```{r ulva-mean-nitrate, fig.height=4.5}
spec_summ_nitrate %>%
  filter(species == "u") %>%
  ggplot(aes(x = wavelength, y = mean_abs, color = nitrate, fill = nitrate)) +
  geom_vline(data = peaks_ulva, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey55", linewidth = 0.4, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - se_abs, ymax = mean_abs + se_abs),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.65) +
  scale_color_viridis_d(option = "plasma", direction = -1) +
  scale_fill_viridis_d(option = "plasma", direction = -1) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean zero-baseline absorptance (SNV)",
    color   = "Nitrate (µmol)",
    fill    = "Nitrate (µmol)",
    title   = expression(italic("Ulva lactuca") ~ "— Mean Absorptance ± SE by Nitrate"),
    caption = "Dashed lines: 435 nm Chl a/b+Car | 500 nm Carotenoid | 572 nm Green trough | 650 nm Chl b | 675 nm Chl a"
  )
```

### *Hypnea musciformis*

```{r hypnea-mean-nitrate, fig.height=4.5}
spec_summ_nitrate %>%
  filter(species == "h") %>%
  ggplot(aes(x = wavelength, y = mean_abs, color = nitrate, fill = nitrate)) +
  geom_vline(data = peaks_hypnea, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey55", linewidth = 0.4, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - se_abs, ymax = mean_abs + se_abs),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.65) +
  scale_color_viridis_d(option = "plasma", direction = -1) +
  scale_fill_viridis_d(option = "plasma", direction = -1) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean zero-baseline absorptance (SNV)",
    color   = "Nitrate (µmol)",
    fill    = "Nitrate (µmol)",
    title   = expression(italic("Hypnea musciformis") ~ "— Mean Absorptance ± SE by Nitrate"),
    caption = "Dashed lines: 435 nm Chl a/b+Car | 500 nm B-PE | 572 nm R-PE | 625 nm PC | 675 nm Chl a"
  )
```

---

## 3. Mean Spectra by Salinity

```{r mean-spectra-salinity}
spec_summ_sal <- spec_norm %>%
  filter(wavelength >= 400, wavelength <= 750) %>%
  group_by(species, salinity, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    se_abs   = sd(absorptance_zero,   na.rm = TRUE) / sqrt(n()),
    .groups  = "drop"
  )
```

```{r ulva-mean-salinity, fig.height=4.5}
spec_summ_sal %>%
  filter(species == "u") %>%
  ggplot(aes(x = wavelength, y = mean_abs, color = salinity, fill = salinity)) +
  geom_vline(data = peaks_ulva, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey55", linewidth = 0.4, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - se_abs, ymax = mean_abs + se_abs),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.65) +
  scale_color_manual(values = c("18" = "#4DBBEE", "22" = "#E8844A")) +
  scale_fill_manual( values = c("18" = "#4DBBEE", "22" = "#E8844A")) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean zero-baseline absorptance (SNV)",
    color   = "Salinity (ppt)",
    fill    = "Salinity (ppt)",
    title   = expression(italic("Ulva lactuca") ~ "— Mean Absorptance ± SE by Salinity"),
    caption = "Dashed lines: 435 nm Chl a/b+Car | 500 nm Carotenoid | 572 nm Green trough | 650 nm Chl b | 675 nm Chl a"
  )
```

```{r hypnea-mean-salinity, fig.height=4.5}
spec_summ_sal %>%
  filter(species == "h") %>%
  ggplot(aes(x = wavelength, y = mean_abs, color = salinity, fill = salinity)) +
  geom_vline(data = peaks_hypnea, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey55", linewidth = 0.4, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - se_abs, ymax = mean_abs + se_abs),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.65) +
  scale_color_manual(values = c("18" = "#4DBBEE", "22" = "#E8844A")) +
  scale_fill_manual( values = c("18" = "#4DBBEE", "22" = "#E8844A")) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean zero-baseline absorptance (SNV)",
    color   = "Salinity (ppt)",
    fill    = "Salinity (ppt)",
    title   = expression(italic("Hypnea musciformis") ~ "— Mean Absorptance ± SE by Salinity"),
    caption = "Dashed lines: 435 nm Chl a/b+Car | 500 nm B-PE | 572 nm R-PE | 625 nm PC | 675 nm Chl a"
  )
```

---

## 4. Mean Spectra by Treatment Group (Nitrate × Salinity)

Eight treatment combinations (4 nitrate × 2 salinity), each plotted as a single mean line.

```{r mean-spectra-combo}
spec_summ_combo <- spec_norm %>%
  filter(wavelength >= 400, wavelength <= 750) %>%
  mutate(group = paste0(nitrate, " µmol / ", salinity, " ppt")) %>%
  group_by(species, nitrate, salinity, group, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(
    nitrate_f  = factor(nitrate,  levels = sort(unique(as.numeric(as.character(nitrate))))),
    salinity_f = factor(salinity, levels = sort(unique(as.numeric(as.character(salinity)))))
  )
```

```{r ulva-combo, fig.height=4.5}
spec_summ_combo %>%
  filter(species == "u") %>%
  ggplot(aes(x = wavelength, y = mean_abs,
             color = nitrate_f, linetype = salinity_f)) +
  geom_vline(data = peaks_ulva, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey70", linewidth = 0.35, inherit.aes = FALSE) +
  geom_line(linewidth = 0.55) +
  scale_color_viridis_d(option = "plasma", direction = -1) +
  scale_linetype_manual(values = c("18" = "solid", "22" = "dashed")) +
  theme_bw() +
  labs(
    x        = "Wavelength (nm)",
    y        = "Mean zero-baseline absorptance (SNV)",
    color    = "Nitrate (µmol)",
    linetype = "Salinity (ppt)",
    title    = expression(italic("Ulva lactuca") ~ "— Mean Spectra by Treatment × Salinity"),
    caption  = "Dashed lines: 435 nm | 500 nm Carotenoid | 572 nm Green trough | 650 nm Chl b | 675 nm Chl a"
  )
```

```{r hypnea-combo, fig.height=4.5}
spec_summ_combo %>%
  filter(species == "h") %>%
  ggplot(aes(x = wavelength, y = mean_abs,
             color = nitrate_f, linetype = salinity_f)) +
  geom_vline(data = peaks_hypnea, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey70", linewidth = 0.35, inherit.aes = FALSE) +
  geom_line(linewidth = 0.55) +
  scale_color_viridis_d(option = "plasma", direction = -1) +
  scale_linetype_manual(values = c("18" = "solid", "22" = "dashed")) +
  theme_bw() +
  labs(
    x        = "Wavelength (nm)",
    y        = "Mean zero-baseline absorptance (SNV)",
    color    = "Nitrate (µmol)",
    linetype = "Salinity (ppt)",
    title    = expression(italic("Hypnea musciformis") ~ "— Mean Spectra by Treatment × Salinity"),
    caption  = "Dashed lines: 435 nm | 500 nm B-PE | 572 nm R-PE | 625 nm PC | 675 nm Chl a"
  )
```

---

## 5. Faceted by Nitrate and Salinity

Individual panels for each nitrate × salinity combination, showing mean ± SD across all individuals in that group.

```{r ulva-faceted, fig.height=6, fig.width=11}
spec_norm %>%
  filter(species == "u", wavelength >= 400, wavelength <= 750) %>%
  group_by(nitrate, salinity, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    sd_abs   = sd(absorptance_zero,   na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  ggplot(aes(x = wavelength, y = mean_abs)) +
  geom_vline(data = peaks_ulva, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey60", linewidth = 0.35, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - sd_abs, ymax = mean_abs + sd_abs),
              fill = "#4dbb8e", alpha = 0.25) +
  geom_line(color = "#1a6b45", linewidth = 0.6) +
  facet_grid(salinity ~ nitrate,
             labeller = labeller(
               salinity = function(x) paste0(x, " ppt"),
               nitrate  = function(x) paste0(x, " µmol NO₃")
             )) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean ± SD absorptance (SNV)",
    title   = expression(italic("Ulva lactuca") ~ "— Absorptance by Nitrate × Salinity"),
    caption = "Ribbon = ±1 SD  |  Dashed lines: 435 | 500 | 572 | 650 nm Chl b | 675 nm Chl a"
  )
```

```{r hypnea-faceted, fig.height=6, fig.width=11}
spec_norm %>%
  filter(species == "h", wavelength >= 400, wavelength <= 750) %>%
  group_by(nitrate, salinity, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    sd_abs   = sd(absorptance_zero,   na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  ggplot(aes(x = wavelength, y = mean_abs)) +
  geom_vline(data = peaks_hypnea, aes(xintercept = wavelength),
             linetype = "dashed", color = "grey60", linewidth = 0.35, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = mean_abs - sd_abs, ymax = mean_abs + sd_abs),
              fill = "#bb4d7a", alpha = 0.25) +
  geom_line(color = "#7a1a45", linewidth = 0.6) +
  facet_grid(salinity ~ nitrate,
             labeller = labeller(
               salinity = function(x) paste0(x, " ppt"),
               nitrate  = function(x) paste0(x, " µmol NO₃")
             )) +
  theme_bw() +
  labs(
    x       = "Wavelength (nm)",
    y       = "Mean ± SD absorptance (SNV)",
    title   = expression(italic("Hypnea musciformis") ~ "— Absorptance by Nitrate × Salinity"),
    caption = "Ribbon = ±1 SD  |  Dashed lines: 435 | 500 nm B-PE | 572 nm R-PE | 625 nm PC | 675 nm Chl a"
  )
```

---

## 6. Both Species Overlaid

Direct comparison of mean absorptance curves for both species across nitrate treatments (salinity pooled).

```{r both-species, fig.height=5}
spec_norm %>%
  filter(wavelength >= 400, wavelength <= 750) %>%
  group_by(species, nitrate, wavelength) %>%
  summarise(
    mean_abs = mean(absorptance_zero, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(species_label = recode(species, u = "Ulva lactuca", h = "Hypnea musciformis")) %>%
  ggplot(aes(x = wavelength, y = mean_abs,
             color = nitrate, linetype = species_label)) +
  geom_vline(xintercept = c(435, 500, 572, 625, 650, 675),
             linetype = "dashed", color = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.6) +
  scale_color_viridis_d(option = "plasma", direction = -1) +
  scale_linetype_manual(values = c("Ulva lactuca" = "solid",
                                   "Hypnea musciformis" = "dashed")) +
  theme_bw() +
  labs(
    x        = "Wavelength (nm)",
    y        = "Mean zero-baseline absorptance (SNV)",
    color    = "Nitrate (µmol)",
    linetype = "Species",
    title    = "Both Species — Mean Absorptance by Nitrate Treatment",
    caption  = "Dashed grey lines: 435 | 500 | 572 | 625 | 650 | 675 nm"
  )
```
