---
title: "SGD and Wastewater Growth of Ulva and Hypnea"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    toc: yes
    toc_float:
      collapsed: true
---
# Run analysis on the combined dataset using 2021 and 2025 data 
### 2021 = SGD, 2025 = Wastewater Scenarios
### Subsetted for analysis of the individual species

```{r, include= FALSE}
#load libraries
#libraries for tidying up
library(gt)
library(purrr)
library(stringr)
library(tidyr)
library(hms)
library(lubridate)
library(ggeffects)

# libraries for analysis
library(lme4)
library(lmerTest)
library(afex)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)

#libraries for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
library(lubridate)
library(paletteer)
```


# Data Wrangling
##Read in combined dataset
```{r}
combined_growth_all <- read.csv("/Users/angela/src/Photosynthesis/data/wastewater/transformed/combined_data_all.csv")
```

### change variables to factors
```{r}
combined_growth_all <- combined_growth_all %>%
  mutate(date = ymd(date)) %>%
  mutate(year = as.factor(year)) %>%
  mutate(salinity = as.factor(salinity)) %>%
  mutate(temp = as.factor(temp)) %>%
  mutate(nitrate = as.factor(nitrate)) %>%
  mutate(run_combo = as.factor(run_combo)) %>%
  mutate(plant_id = as.factor(plant_id))

```

### Filter data before proceeding -- remove 30 degrees, keep 80 nitrate from SGD 2021 
### Keep all of 2025
```{r}
combined_growth_all <- combined_growth_all %>%
  filter(
    # keep:
    (
      year %in% c(2021, 2022) & 
        nitrate == "80" & 
        temp != 30
    ) |
      year == 2025
  )  
```

### Make a list to be used for plots that orders chronologically first then by nitrate value low to high
```{r}
combined_growth_all$treatment_graph[combined_growth_all$treat_letter == "f"] <- "i) 80 μmol"
combined_growth_all$treatment_graph[combined_growth_all$treat_letter == "k"] <- "k) 80 μmol"
combined_growth_all$treatment_graph[combined_growth_all$treat_letter == "l"] <- "l) 245 μmol"
combined_growth_all$treatment_graph[combined_growth_all$treat_letter == "m"] <- "m) 748 μmol"
combined_growth_all$treatment_graph[combined_growth_all$treat_letter == "n"] <- "n) 2287 μmol"
```

### Subset by species and run table checks
```{r}
combined_growth_u <- combined_growth_all %>%
  filter(species == "u")

combined_growth_h <- combined_growth_all %>%
  filter(species == "h")

xtabs(~ salinity + temp + year, data = combined_growth_u)
xtabs(~ salinity + temp + year, data = combined_growth_h)
```

# Run lmm model for Ulva______________________________________________________
#### Note: All other random effects had zero effect on Ulva growth
```{r}
growth_model_u <- lmer(formula = growth_d9 ~ nitrate + salinity + temp +
                         (1 | run_combo), data = combined_growth_u, REML = TRUE)
```

### Check the performance of the model for dataset: u
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hist(resid(growth_model_u))
performance::check_model(growth_model_u)
check_collinearity(growth_model_u)
r.squaredGLMM(growth_model_u)
summary(growth_model_u)
```

### View model outcome and random effects levels
```{r, fig.width = 8, fig.height = 9, results ='markup'}
tab_model(growth_model_u, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, 
          show.df = TRUE, show.zeroinf = TRUE)
```

## Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
# Does nitrate affect Ulva growth?
u_growth_nitrate_null <- lmer(formula = growth_d9 ~ temp + salinity +
                              (1 | run_combo), 
                              data = combined_growth_u, REML = FALSE)
u_growth_model2 <- lmer(formula = growth_d9 ~ nitrate + temp + salinity +
                          (1 | run_combo), 
                        data = combined_growth_u, REML = FALSE)
anova(u_growth_nitrate_null, u_growth_model2)

```

```{r}
# Does temperature affect Ulva growth?
u_growth_temp_null <- lmer(formula = growth_d9 ~ nitrate + salinity +
                             (1 | run_combo), 
                           data = combined_growth_u, REML = FALSE)
u_growth_model4 <- lmer(formula = growth_d9 ~ nitrate + temp + salinity +
                          (1 | run_combo), 
                        data = combined_growth_u, REML = FALSE)
anova(u_growth_temp_null, u_growth_model4)

```

```{r}
# Does salinity affect Ulva growth?
u_growth_salinity_null <- lmer(formula = growth_d9 ~ nitrate + temp +
                             (1 | run_combo), 
                           data = combined_growth_u, REML = FALSE)
u_growth_model4 <- lmer(formula = growth_d9 ~ nitrate + temp + salinity +
                          (1 | run_combo), 
                        data = combined_growth_u, REML = FALSE)
anova(u_growth_salinity_null, u_growth_model4)
```

# Run lmm model for Hypnea____________________________________________________
### These random effects have been tested and found relevant
```{r}
growth_model_h <- lmer(formula = growth_d9 ~ nitrate + temp + salinity +
                         (1 | plant_id) + (1 | run_combo) + (1 | illumination), 
                       data = combined_growth_h, REML = TRUE)
```

### Check the performance of the model for dataset: h
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hist(resid(growth_model_h))
performance::check_model(growth_model_h)
r.squaredGLMM(growth_model_h)
summary(growth_model_h)
```

### View random effects levels
```{r, fig.width = 8, fig.height = 9, results ='markup'}
tab_model(growth_model_h, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, 
          show.df = TRUE, show.zeroinf = TRUE)
```

## Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
# Does nitrate affect Hypnea growth?
h_growth_nitrate_null <- lmer(formula = growth_d9 ~ salinity + temp + 
                                (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                              data = combined_growth_h, REML = FALSE)
h_growth_model2 <- lmer(formula = growth_d9 ~ nitrate + salinity + temp + 
                          (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                        data = combined_growth_h, REML = FALSE)
anova(h_growth_nitrate_null, h_growth_model2)
```

```{r}
#Does salinity affect Hypnea growth?
h_growth_salinity_null <- lmer(formula = growth_d9 ~ nitrate + temp + 
                                 (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                               data = combined_growth_h, REML = FALSE)
h_growth_model3 <- lmer(formula = growth_d9 ~ nitrate + salinity + temp + 
                          (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                        data = combined_growth_h, REML = FALSE)
anova(h_growth_salinity_null, h_growth_model3)

```

```{r}
h_growth_temp_null <- lmer(formula = growth_d9 ~ nitrate + salinity + 
                             (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                           data = combined_growth_h, REML = FALSE)
h_growth_model4 <- lmer(formula = growth_d9 ~ nitrate + salinity + temp + 
                          (1 | run_combo) + (1 | plant_id) + (1 | illumination), 
                        data = combined_growth_h, REML = FALSE)
anova(h_growth_temp_null, h_growth_model4)
```

# MAKE PLOTS_____________________________________________________________________

### All data included in this plot
```{r, fig.width = 8, fig.height = 9, results ='markup'}
combo_growth_species_plot <- combined_growth_all %>% 
  ggplot(aes(treatment_graph, growth_d9, color = salinity, shape = temp)) + 
  geom_boxplot(linewidth=0.25) + 
  geom_point(alpha = 0.75,
             size = 2,
             position = position_jitter(width = 0.3), show.legend = TRUE) + 
  labs(x="nitrate (μmols)", y= "9-Day Growth (%)", title = "By Species") + 
  facet_wrap(~species, labeller = labeller(species = c("h" = "Hypnea", "u" = "Ulva"))) +
  scale_x_discrete(labels = c("80", "80", "245", "748", "2287")) + 
  ylim(-50, 380) + stat_mean() + 
  geom_hline(yintercept=0, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  scale_color_paletteer_d("MetBrewer::Cross") +
  theme(plot.title = element_text(face = "bold", vjust = -25, hjust = 0.05), # By Species title
        legend.position = c(0.40,0.82), #move it to center of plot
        legend.box = "horizontal", #make the two legend boxes stay side-by-side
        legend.text = element_text(size = 12), #increase size for legibility
        strip.text = element_text(face = "italic", size = 12), #increase size of Hypnea and Ulva in facet headers
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14))
combo_growth_species_plot
```

