---
title: "LMM for Wastewater and SGD combined datasets Hsat and DSPI"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    toc: yes
    toc_float:
      collapsed: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```
### This script has been re-written using tidyverse and its ilk. New functions and structure written to run model on photosynthesis data for Hypnea and Ulva wastewater data PLUS some of the overlapping treatments from the SGD runs in 2021-2023
### This script pulls in the combined dataset for the 2021, 2023 SGD and 2025 wastewater data and keeps all of 2025, and only the 80 nitrate treatment from 2021-2022, which overlaps with 2025.
### It also removes the 30 degree temp treatment from 2021-2022, which is not present in 2025.
### sal_bin predictor (35 ppt = 1, 28, 27 ppt = 2, 18-22 ppt = 3, 11-15 ppt = 4)

#______________________________________
### Load libraries

```{r}
# data wrangling
library(tidyverse) #includes ggplot2, readr, forcats, tibble, purrr, tidyr, dplyr, stringr
library(hms)
library(lubridate)
# mixed models
library(lme4)
library(lmerTest)
library(MuMIn)
library(emmeans)
library(multcomp)
library(ggeffects)
library(performance)
library(ggpubr)
library(sjPlot)
# optional for plots
library(RColorBrewer)
```

### Load this file for normalized to quantum efficiency of photosynthesis per Silsbe and Kromkamp

```{r}
ww_hsat_lmm <- read_csv("/Users/angela/src/Photosynthesis/data/wastewater/transformed/ww_combo_hsat_dspi.csv")
```

### Remove individual that died during experiment

```{r}
ww_hsat_lmm <- ww_hsat_lmm %>%
  filter(id != "ul09")
```

### Format date and time correctly
```{r}
TZ <- "Pacific/Honolulu"
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(
    start_date = as.Date(start_date),
    start_time = hms::as_hms(start_time),
    end_time   = hms::as_hms(end_time)
  )
```

### Make unique_id column for troubleshooting individuals

```{r}
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(unique_id = paste(start_date, id, sep = "_"))
```

```{r}
# assign list of variables as factor
factor_vars <- c("sal_bin", "temp", "nitrate", "run_combo", "salinity")
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(across(all_of(factor_vars), as.factor))
```

### Assign new column for chronological ranking of individuals in each day_group
### Use ranking to make smaller groups of ~15 minutes for rlc_order
```{r}
ww_hsat_lmm <- ww_hsat_lmm %>%
  group_by(start_date, rlc_day) %>%
  arrange(end_time, .by_group = TRUE) %>%
  mutate(
    rank = dplyr::row_number(),                 # stable rank by time
    rlc_order1 = floor((rank - 1) / 3) + 1      # 3 steps ≈ 15 minutes
  ) %>%
  ungroup()
```

### Add a year column for subsetting later
```{r}
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(year = lubridate::year(start_date))
```


### Convert DSPI from micromols to mols m-2 (dspi_day_mean computed below after run_days is known)
```{r}
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(
    dspi_mol = round(dspi_total / 1e6, 2)
  )
ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(
    start_dt = lubridate::as_datetime(start_date, tz = TZ) + start_time,
    end_dt   = lubridate::as_datetime(start_date + days(8), tz = TZ) + end_time,  # all experiments ran day 1–9, so 8 elapsed days
    run_hours = as.numeric(difftime(end_dt, start_dt, units = "hours")),
    run_days  = run_hours / 24
  ) %>%
  mutate(
    # mean per day + round to 2 decimals
    hsat_day_mean = round(hsat_total / run_days, 2),
    relhsat_day_mean = round(relhsat_total / run_days, 2),
    dspi_day_mean = round(dspi_mol / run_days, 2)
  )
```

### Check data before moving on
```{r}
ww_hsat_lmm %>%
  summarise(
    min_days = min(run_days, na.rm = TRUE),
    max_days = max(run_days, na.rm = TRUE),
    n_bad = sum(is.na(run_days) | run_days <= 0, na.rm = TRUE)
  )
```

### Join growth rates from combined_data_all
### run_combo numbering differs between files so cannot be used as join key.
### id alone is also not unique (reused across years and runs within the same year).
### The only collision-free key is id + end_date, where end_date = start_date + 8 days,
### which matches the 'date' column in combined_data_all.
```{r}
growth_key <- read.csv("/Users/angela/src/Photosynthesis/data/wastewater/transformed/combined_data_all.csv") %>%
  dplyr::select(id, date, growth_d9) %>%
  mutate(id       = tolower(trimws(id)),
         end_date = as.Date(date))

ww_hsat_lmm <- ww_hsat_lmm %>%
  mutate(end_date = as.Date(start_date) + days(8)) %>%
  left_join(growth_key, by = c("id", "end_date")) %>%
  mutate(run_combo = as.factor(run_combo))
```

### New dataset with overlaps from 2021, 2022 and all of 2025
### This keeps all of 2025, and only the 80 nitrate treatment from 2021-2022.
### It also removes the 30 degree temp treatment from 2021-2022, which is not present in 2025.

```{r}
sgd_ww_hsat <- ww_hsat_lmm %>%
  filter(
    # keep:
    (
      year %in% c(2021, 2022) &
        nitrate == "80" &
        temp != 30
    ) |
      year == 2025
  )
```

### Subset the species
```{r}
ulva_ww_hsat <- subset(sgd_ww_hsat, species == "u")

hypnea_ww_hsat <- subset(sgd_ww_hsat, species == "h")

ulva_2025   <- subset(ww_hsat_lmm, species == "u" & year == 2025)
hypnea_2025 <- subset(ww_hsat_lmm, species == "h" & year == 2025)
```

### Check data structure
```{r}
xtabs(~ salinity + temp + year, data = ulva_ww_hsat)
xtabs(~ salinity + temp + year, data = hypnea_ww_hsat)
```

# MODEL FUNCTION__________________________________________________________________

```{r}
check_model_fit <- function(model) {
  hist(resid(model))
  plot(resid(model) ~ fitted(model))
  qqnorm(resid(model))
  qqline(resid(model))
  print(performance::check_model(model))
  print(MuMIn::r.squaredGLMM(model))
}
```

### Get predictor means
```{r}
get_data_means <- function(data, predictors_list, response){
  data %>%
    group_by(across(all_of(predictors_list))) %>%
    summarise(!!paste0(response, "_mean") := mean(.data[[response]], na.rm = TRUE),
              .groups = "drop")
}
```

### Make a function to run the models for likelihood ratio tests where REML must be FALSE
### Construct null model to perform likelihood ratio test. REML must be FALSE.
### Optimizer set to bobyqa with more iterations to suppress non-convergence warnings
### common with small numbers of random effect levels.

```{r}
compare_lmer_models <- function(data, response, predictors_list, random_effects) {
  # remove NA/NaN/Inf ONLY for the response variable
  data_clean <- data %>%
    filter(is.finite(.data[[response]])) %>%              # drop NA / NaN / Inf in response
    tidyr::drop_na(all_of(predictors_list)) %>%           # drop rows missing predictor values
    droplevels()

  # Build random-effects string
  re_str <- paste(paste0("(1 | ", random_effects, ")"), collapse = " + ")

  # Optimizer control: bobyqa with more iterations suppresses non-convergence
  # warnings common with small numbers of random effect levels
  lmer_ctrl <- lme4::lmerControl(optimizer = "bobyqa",
                                 optCtrl   = list(maxfun = 2e5))

  # Full model (REML = FALSE required for LRTs)
  fixed_all <- paste(predictors_list, collapse = " + ")
  f_full <- as.formula(paste(response, "~", fixed_all, "+", re_str))
  model_all <- lme4::lmer(f_full, data = data_clean, REML = FALSE, control = lmer_ctrl)

  # Refit full model with REML = TRUE for reporting coefficients and diagnostics
  model_all_reml <- lme4::lmer(f_full, data = data_clean, REML = TRUE, control = lmer_ctrl)

  # Drop-one models + LRTs
  models <- list()
  anova_result <- list()

  for (dropped in predictors_list) {
    fixed_reduced <- paste(setdiff(predictors_list, dropped), collapse = " + ")
    f_reduced <- as.formula(paste(response, "~", fixed_reduced, "+", re_str))
    models[[dropped]] <- lme4::lmer(f_reduced, data = data_clean, REML = FALSE,
                                    control = lmer_ctrl)
    anova_result[[dropped]] <- anova(models[[dropped]], model_all, test = "Chisq")
  }

  data_means <- get_data_means(data_clean, predictors_list, response)

  list(
    models = models,          # named by the predictor that was dropped
    model_all = model_all,    # ML fit — use only for LRTs
    model_all_reml = model_all_reml,  # REML fit — use for reporting and diagnostics
    anova_result = anova_result,
    data_means = data_means,
    used_data = data_clean
  )
}

predictors_list         <- c("nitrate", "temp")
predictors_list_species <- c("nitrate", "temp", "species")
```

### Model selection note — combined 2021+2022 and 2025 dataset
Full fixed effects tested: salinity + nitrate + temp. Salinity was dropped from all
combined-dataset models: including it caused rank deficiency in the fixed-effect model
matrix across every response variable and both species (confirmed by "fixed-effect model
matrix is rank deficient" warning). Rank deficiency arises because salinity treatment
structure is collinear with temperature across the two datasets — certain
salinity-temperature combinations exist only in 2021 or only in 2025, making their
effects inseparable. For DSPI models, rlc_order1 was also dropped from random effects
to resolve singular fit. Final fixed effects retained: nitrate + temp. Salinity IS
included as a predictor in the 2025-only models below.

# Hsat_____________________________________
### Inputs for Ulva/Hsat
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_ww_hsat_results <- compare_lmer_models(
  data = ulva_ww_hsat,
  response = "hsat_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_hsat_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Ulva hsat?
ulva_ww_hsat_results$anova_result[["nitrate"]]
# Does temp affect Ulva hsat?
ulva_ww_hsat_results$anova_result[["temp"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ulva_ww_hsat_results$model_all_reml)
ulva_ww_hsat_results$data_means
```

### Inputs for Hypnea/Hsat
```{r}
hypnea_ww_hsat_results <- compare_lmer_models(
  data = hypnea_ww_hsat,
  response = "hsat_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_hsat_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Hypnea hsat?
hypnea_ww_hsat_results$anova_result[["nitrate"]]
# Does temp affect Hypnea hsat?
hypnea_ww_hsat_results$anova_result[["temp"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(hypnea_ww_hsat_results$model_all_reml)
hypnea_ww_hsat_results$data_means
```

## Compare species
### Inputs for Hsat species comparison
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ww_hsat_species_results <- compare_lmer_models(
  data = sgd_ww_hsat,
  response = "hsat_day_mean",
  predictors_list_species,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_hsat_species_results$model_all_reml)
```
### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does species affect hsat?
ww_hsat_species_results$anova_result[["species"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ww_hsat_species_results$model_all_reml)
ww_hsat_species_results$data_means
```

### relHsat____________________________________

### Inputs for Ulva relHsat
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_ww_relhsat_results <- compare_lmer_models(
  data = ulva_ww_hsat,
  response = "relhsat_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_relhsat_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Ulva relhsat?
ulva_ww_relhsat_results$anova_result[["nitrate"]]
# Does temp affect Ulva relhsat?
ulva_ww_relhsat_results$anova_result[["temp"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ulva_ww_relhsat_results$model_all_reml)
ulva_ww_relhsat_results$data_means
```

### Inputs for Hypnea relHsat
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hypnea_ww_relhsat_results <- compare_lmer_models(
  data = hypnea_ww_hsat,
  response = "relhsat_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_relhsat_results$model_all_reml)
```
### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Hypnea relhsat?
hypnea_ww_relhsat_results$anova_result[["nitrate"]]
# Does temp affect Hypnea relhsat?
hypnea_ww_relhsat_results$anova_result[["temp"]]
```
### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(hypnea_ww_relhsat_results$model_all_reml)
hypnea_ww_relhsat_results$data_means
```

### Inputs for relHsat species comparison
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ww_relhsat_species_results <- compare_lmer_models(
  data = sgd_ww_hsat,
  response = "relhsat_day_mean",
  predictors_list_species,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_relhsat_species_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does species affect relhsat?
ww_relhsat_species_results$anova_result[["species"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ww_relhsat_species_results$model_all_reml)
ww_relhsat_species_results$data_means
```

### DSPI_____________________________________
### Inputs for Ulva DSPI
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_ww_dspi_results <- compare_lmer_models(
  data = ulva_ww_hsat,
  response = "dspi_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_dspi_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Ulva dspi?
ulva_ww_dspi_results$anova_result[["nitrate"]]
# Does temp affect Ulva dspi?
ulva_ww_dspi_results$anova_result[["temp"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ulva_ww_dspi_results$model_all_reml)
ulva_ww_dspi_results$data_means
```

### Inputs for Hypnea/DSPI
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hypnea_ww_dspi_results <- compare_lmer_models(
  data = hypnea_ww_hsat,
  response = "dspi_day_mean",
  predictors_list,
  random_effects = c("plant_id", "run_combo")
)
```

### Access results
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_dspi_results$model_all_reml)
```

### Run ANOVAs
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does nitrate affect Hypnea dspi?
hypnea_ww_dspi_results$anova_result[["nitrate"]]
# Does temp affect Hypnea dspi?
hypnea_ww_dspi_results$anova_result[["temp"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(hypnea_ww_dspi_results$model_all_reml)
hypnea_ww_dspi_results$data_means
```

### Inputs for DSPI species comparison
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ww_dspi_species_results <- compare_lmer_models(
  data = sgd_ww_hsat,
  response = "dspi_day_mean",
  predictors_list_species,
  random_effects = c("plant_id", "run_combo")
)
```

### Access results — Full model
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_dspi_species_results$model_all_reml)
```

### Run ANOVA
```{r, fig.width = 8, fig.height = 9, results ='markup'}
# Does species affect dspi?
ww_dspi_species_results$anova_result[["species"]]
```

### Check model fit and get means table
```{r, fig.width = 8, fig.height = 9, results ='markup'}
check_model_fit(ww_dspi_species_results$model_all_reml)
ww_dspi_species_results$data_means
```

## Model Summary Tables_____________________________________________________

### This produces individual outputs for each model
```{r}
print_model <- function(model, species, response) {
  sjPlot::tab_model(
    model,
    title = paste(species, "-", response),
    show.intercept = TRUE,
    show.se = TRUE,
    show.stat = TRUE,
    show.df = TRUE
  )
}
print_model(ulva_ww_hsat_results$model_all_reml, "Ulva", "Hsat day mean")
print_model(hypnea_ww_hsat_results$model_all_reml, "Hypnea", "Hsat day mean")

print_model(ulva_ww_relhsat_results$model_all_reml, "Ulva", "Relative Hsat")
print_model(hypnea_ww_relhsat_results$model_all_reml, "Hypnea", "Relative Hsat")

print_model(ulva_ww_dspi_results$model_all_reml, "Ulva", "DSPI day mean")
print_model(hypnea_ww_dspi_results$model_all_reml, "Hypnea", "DSPI day mean")
```

## HISTOGRAMS and PLOTS_____________________________________
### Function for raw data plots
```{r}
raw_plots <- function(data, response, response2, label, pretty_color, aescolor, x, x2, y, y2, title, title2, subtitle, ylim1, ylim2,
                      color1, color2, color3, yint, vjust_t, hjust_t, vjust_s, hjust_s, labels1) {
  histo <- data %>%
    ggplot(aes(x = {{response}})) +
    geom_histogram(fill = pretty_color, color = "black", linewidth = 0.25, alpha = 0.85) +
    labs(title = label) +
    theme_bw()

  plot <- data %>%
    ggplot(aes(x = nitrate, y = {{response}})) +
    geom_boxplot(size=0.5) +
    geom_point(alpha = 0.75, size = 3, aes(color = {{aescolor}}), position = "jitter", show.legend = TRUE) +
    labs(x = x, y = y, title = title, subtitle = subtitle) +
    ylim(ylim1, ylim2) + stat_mean() +
    scale_color_manual(values = c(color1, color2, color3)) +
    scale_x_discrete(labels = labels1) +
    geom_hline(yintercept=yint, color = "red", linewidth = 0.5, alpha = 0.5) +
    theme_bw() +
    theme(legend.position = c(0.90,0.80), plot.title = element_text(face = "italic", vjust = vjust_t, hjust = hjust_t),
          plot.subtitle = element_text(face = "bold", size = 14, vjust = vjust_s, hjust = hjust_s))

  return(list(
    histo = histo,
    plot = plot))
}
```


## INPUTS for histo and plots________________________

### Hsat ------------------------

### Inputs for Ulva/Hsat
```{r}
hsat_ulva <- raw_plots(
  data = ulva_ww_hsat,
  response = hsat_day_mean,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "goldenrod1",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Hsat (min/day)",
  y = "Day 9 Hsat (min/day)",
  y2 = "9-Day Growth (%)",
  title = "A - Ulva lactuca",
  title2 = "Ulva lactuca --- Hsat (min/day) vs 9-Day Growth (%)",
  subtitle = "Hsat",
  ylim1 = 100,
  ylim2 = 1000,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 400,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol",
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(hsat_ulva$histo)
plot(hsat_ulva$plot)
```

### Inputs for Hypnea/Hsat
```{r}
hsat_hypnea <- raw_plots(
  data = hypnea_ww_hsat,
  response = hsat_day_mean,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon4",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Hsat (min/day)",
  y = "Day 9 Hsat (min/day)",
  y2 = "9-Day Growth (%)",
  title = "B - Hypnea musciformis",
  title2 = "Hypnea musciformis --- Hsat vs 9-Day Growth (%)",
  subtitle = "Hsat",
  ylim1 = 100,
  ylim2 = 1000,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 20,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol",
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(hsat_hypnea$histo)
plot(hsat_hypnea$plot)
```

### DSPI ------------------------

### Inputs for Ulva/DSPI
```{r}
dspi_ulva <- raw_plots(
  data = ulva_ww_hsat,
  response = dspi_day_mean,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "goldenrod1",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "DSPI (mol/day)",
  y = "Day 9 DSPI (mol/day)",
  y2 = "9-Day Growth (%)",
  title = "A - Ulva lactuca",
  title2 = "Ulva lactuca --- DSPI (mol/day) vs 9-Day Growth (%)",
  subtitle = "DSPI",
  ylim1 = 0,
  ylim2 = 5,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 2,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol",
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(dspi_ulva$histo)
plot(dspi_ulva$plot)
```

### Inputs for Hypnea/DSPI
```{r}
dspi_hypnea <- raw_plots(
  data = hypnea_ww_hsat,
  response = dspi_day_mean,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon4",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "DSPI (mol/day)",
  y = "Day 9 DSPI (mol/day)",
  y2 = "9-Day Growth (%)",
  title = "B - Hypnea musciformis",
  title2 = "Hypnea musciformis --- DSPI vs 9-Day Growth (%)",
  subtitle = "DSPI",
  ylim1 = 0,
  ylim2 = 5,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 2,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol",
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(dspi_hypnea$histo)
plot(dspi_hypnea$plot)
```

## DSPI vs Growth linear regression ____________________________________________
### DSPI (mean daily, mol O2 m-2 day-1) as predictor of 9-day growth (%)
### Points coloured by nitrate treatment; single regression line across all treatments.
### Annotation shows equation with slope, R2, and p-value.

```{r, fig.width = 6, fig.height = 5}
dspi_growth_ulva <- ulva_ww_hsat %>%
  filter(!is.na(dspi_day_mean), !is.na(growth_d9)) %>%
  ggplot(aes(x = dspi_day_mean, y = growth_d9)) +
  geom_point(aes(color = nitrate), alpha = 0.75, size = 2.5) +
  geom_smooth(method = "lm", color = "black", linewidth = 0.8, se = TRUE) +
  stat_regline_equation(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
    label.x.npc = "left", label.y.npc = 0.97, size = 3.8
  ) +
  stat_cor(
    aes(label = ..p.label..),
    label.x.npc = "left", label.y.npc = 0.90, size = 3.8
  ) +
  scale_color_viridis_d(option = "viridis", direction = -1) +
  labs(
    x     = "Mean daily DSPI (mol O2 m-2 day-1)",
    y     = "9-day growth (%)",
    title = "Ulva lactuca",
    color = "Nitrate (\u00b5mol)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(face = "italic", size = 13),
        axis.title = element_text(size = 12))
plot(dspi_growth_ulva)
```

```{r, fig.width = 6, fig.height = 5}
dspi_growth_hypnea <- hypnea_ww_hsat %>%
  filter(!is.na(dspi_day_mean), !is.na(growth_d9)) %>%
  ggplot(aes(x = dspi_day_mean, y = growth_d9)) +
  geom_point(aes(color = nitrate), alpha = 0.75, size = 2.5) +
  geom_smooth(method = "lm", color = "black", linewidth = 0.8, se = TRUE) +
  stat_regline_equation(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
    label.x.npc = "left", label.y.npc = 0.97, size = 3.8
  ) +
  stat_cor(
    aes(label = ..p.label..),
    label.x.npc = "left", label.y.npc = 0.90, size = 3.8
  ) +
  scale_color_viridis_d(option = "magma", direction = -1) +
  labs(
    x     = "Mean daily DSPI (mol O2 m-2 day-1)",
    y     = "9-day growth (%)",
    title = "Hypnea musciformis",
    color = "Nitrate (\u00b5mol)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(face = "italic", size = 13),
        axis.title = element_text(size = 12))
plot(dspi_growth_hypnea)
```

### Both species side by side
```{r, fig.width = 11, fig.height = 5}
dspi_growth_both <- ggarrange(
  dspi_growth_ulva, dspi_growth_hypnea,
  ncol = 2, nrow = 1,
  labels = c("A", "B")
)
plot(dspi_growth_both)
```

# 2025-ONLY COMPARISON MODELS ________________________________________________
### Predictors: nitrate + salinity only (temp is constant at 22C in 2025)
### All 2025 data is at 22 degrees C so temp cannot be used as a predictor.
### plant_id and rlc_order1 collapse to zero variance in 2025 DSPI models;
### retain run_combo only for DSPI. Hsat retains all three random effects.

```{r}
predictors_list_2025         <- c("nitrate", "salinity")
predictors_list_2025_species <- c("nitrate", "salinity", "species")
```

## Hsat — 2025 only

### Ulva Hsat 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_2025_hsat_results <- compare_lmer_models(
  data = ulva_2025,
  response = "hsat_day_mean",
  predictors_list_2025,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
summary(ulva_2025_hsat_results$model_all_reml)
ulva_2025_hsat_results$anova_result[["nitrate"]]
ulva_2025_hsat_results$anova_result[["salinity"]]
check_model_fit(ulva_2025_hsat_results$model_all_reml)
ulva_2025_hsat_results$data_means
```

### Hypnea Hsat 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hypnea_2025_hsat_results <- compare_lmer_models(
  data = hypnea_2025,
  response = "hsat_day_mean",
  predictors_list_2025,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
summary(hypnea_2025_hsat_results$model_all_reml)
hypnea_2025_hsat_results$anova_result[["nitrate"]]
hypnea_2025_hsat_results$anova_result[["salinity"]]
check_model_fit(hypnea_2025_hsat_results$model_all_reml)
hypnea_2025_hsat_results$data_means
```

## relHsat — 2025 only

### Ulva relHsat 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_2025_relhsat_results <- compare_lmer_models(
  data = ulva_2025,
  response = "relhsat_day_mean",
  predictors_list_2025,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
summary(ulva_2025_relhsat_results$model_all_reml)
ulva_2025_relhsat_results$anova_result[["nitrate"]]
ulva_2025_relhsat_results$anova_result[["salinity"]]
check_model_fit(ulva_2025_relhsat_results$model_all_reml)
ulva_2025_relhsat_results$data_means
```

### Hypnea relHsat 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hypnea_2025_relhsat_results <- compare_lmer_models(
  data = hypnea_2025,
  response = "relhsat_day_mean",
  predictors_list_2025,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
summary(hypnea_2025_relhsat_results$model_all_reml)
hypnea_2025_relhsat_results$anova_result[["nitrate"]]
hypnea_2025_relhsat_results$anova_result[["salinity"]]
check_model_fit(hypnea_2025_relhsat_results$model_all_reml)
hypnea_2025_relhsat_results$data_means
```

## DSPI — 2025 only
### plant_id and rlc_order1 collapse to zero for DSPI; run_combo only retained.

### Ulva DSPI 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
ulva_2025_dspi_results <- compare_lmer_models(
  data = ulva_2025,
  response = "dspi_day_mean",
  predictors_list_2025,
  random_effects = c("run_combo")
)
summary(ulva_2025_dspi_results$model_all_reml)
ulva_2025_dspi_results$anova_result[["nitrate"]]
ulva_2025_dspi_results$anova_result[["salinity"]]
check_model_fit(ulva_2025_dspi_results$model_all_reml)
ulva_2025_dspi_results$data_means
```

### Hypnea DSPI 2025
```{r, fig.width = 8, fig.height = 9, results ='markup'}
hypnea_2025_dspi_results <- compare_lmer_models(
  data = hypnea_2025,
  response = "dspi_day_mean",
  predictors_list_2025,
  random_effects = c("run_combo")
)
summary(hypnea_2025_dspi_results$model_all_reml)
hypnea_2025_dspi_results$anova_result[["nitrate"]]
hypnea_2025_dspi_results$anova_result[["salinity"]]
check_model_fit(hypnea_2025_dspi_results$model_all_reml)
hypnea_2025_dspi_results$data_means
```

## Species comparisons — 2025 only

```{r, fig.width = 8, fig.height = 9, results ='markup'}
ww_2025_hsat_species_results <- compare_lmer_models(
  data = subset(ww_hsat_lmm, year == 2025),
  response = "hsat_day_mean",
  predictors_list_2025_species,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
ww_2025_hsat_species_results$anova_result[["species"]]

ww_2025_dspi_species_results <- compare_lmer_models(
  data = subset(ww_hsat_lmm, year == 2025),
  response = "dspi_day_mean",
  predictors_list_2025_species,
  random_effects = c("run_combo")
)
ww_2025_dspi_species_results$anova_result[["species"]]
```

## Model summary tables — 2025 only

```{r}
print_model(ulva_2025_hsat_results$model_all_reml,   "Ulva 2025",   "Hsat day mean")
print_model(hypnea_2025_hsat_results$model_all_reml, "Hypnea 2025", "Hsat day mean")
print_model(ulva_2025_dspi_results$model_all_reml,   "Ulva 2025",   "DSPI day mean")
print_model(hypnea_2025_dspi_results$model_all_reml, "Hypnea 2025", "DSPI day mean")
```
