---
title: "LMM for Wastewater and SGD combined datasets Pmax, Ek, NPQ"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    toc: yes
    toc_float:
      collapsed: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

### This script has been re-written using tidyverse and its ilk. New functions and structure written to run model on photosynthesis data for Hypnea and Ulva wastewater data PLUS some of the overlapping treatments from the SGD runs in 2021-2023
### This script pulls in the combined dataset for the 2021, 2023 SGD and 2025 wastewater data and keeps all of 2025, and only the 53 and 80 nitrate treatments from 2021-2023, which are the only ones that overlap with 2025.
### It also removes the 30 degree temp treatment from 2021-2023, which is not present in 2025.
### sal_bin predictor (35 ppt = 1, 28, 27 ppt = 2, 18-22 ppt = 3, 11-15 ppt = 4)

#______________________________________
### load libraries

```{r}
# data wrangling
library(tidyverse) #includes ggplot2, readr, forcats, tibble, purrr, tidyr, dplyr, stringr
library(hms)
library(lubridate)
# mixed effect models
library(lme4)
library(lmerTest)
library(MuMIn)
library(emmeans)
library(multcomp)
library(ggeffects)
library(performance)
library(ggpubr)
# optional for plots
library(RColorBrewer)
```

### Load this file (data normalized to quantum efficiency of photosynthesis per Silsbe and Kromkamp)
```{r}
ww_lmm <- read_csv("/Users/angela/src/Photosynthesis/data/wastewater/transformed/ww_combo_ek_alpha_norm.csv")

glimpse(ww_lmm)

ww_lmm <- ww_lmm %>%
  mutate(
    date = as.Date(date),
    rlc_time = as_hms(rlc_time),
    rlc_end_time = as_hms(rlc_end_time)
  )
```

### Assign list of variables as factor
```{r}
factor_vars <- c("treat_letter", "treatment", "salinity", 
"sal_bin", "temp", "rlc_day", "nitrate", "run_combo")

ww_lmm <- ww_lmm %>%
  mutate(across(all_of(factor_vars), as.factor))
```

### Assign new column for chronological ranking of individuals in each day_group
### Use ranking to make smaller groups of ~15 minutes for rlc_order
```{r}
ww_lmm <- ww_lmm %>%
  group_by(date, rlc_day) %>%
  arrange(rlc_end_time, .by_group = TRUE) %>%
  mutate(
    rank = dplyr::row_number(),                 # stable rank by time
    rlc_order1 = floor((rank - 1) / 3) + 1      # 3 steps ≈ 15 minutes
  ) %>%
  ungroup()
```

### Treatment Graph is set up to have predictors in order
```{r}
ww_lmm <- ww_lmm %>%
  mutate(treatment_graph = case_when(
    treat_letter == "a" ~ "1) 0.5 μmol",
    treat_letter == "b" ~ "3) 14 μmol",
    treat_letter == "c" ~ "4) 27 μmol",
    treat_letter == "d" ~ "6) 53 μmol",
    treat_letter == "e" ~ "7) 53 μmol",
    treat_letter == "f" ~ "9) 80 μmol",
    treat_letter == "g" ~ "2) 0.5 μmol",
    treat_letter == "h" ~ "5) 37 μmol",
    treat_letter == "i" ~ "8) 53 μmol",
    treat_letter == "j" ~ "10) 86 μmol",
    treat_letter == "k" ~ "11) 80 μmol",
    treat_letter == "l" ~ "12) 245 μmol",
    treat_letter == "m" ~ "13) 748 μmol",
    treat_letter == "n" ~ "14) 2287 μmol"
  ))
```

### Clean dates to be sure
```{r}
ww_lmm <- ww_lmm %>%
  mutate(year = lubridate::year(date))
```

### Subset for overlapping treatments, remove 30 degree treatment from 2021
### New subset keeps all 2025 wastewater and 80 umol N from 2021
```{r}
sgd_ww <- ww_lmm %>%
  filter(
    # keep:
    (
      year %in% c(2021, 2022) & 
        nitrate == "80"   ) |
      year == 2025
  )
```

### Subset for day 9 only and then subset for each species
```{r}
ww_lmm_d9 <- sgd_ww %>%
  filter(rlc_day == 9) #only d9 is included in dataset
# subset the species
ulva_ww <- subset(ww_lmm_d9, species == "u")

hypnea_ww <- subset(ww_lmm_d9, species == "h")
```

# Run the models -- start with the performance checks and means
```{r}
check_model_fit <- function(model, terms) {
  hist(resid(model))
  plot(resid(model) ~ fitted(model))
  qqnorm(resid(model))
  qqline(resid(model))
  print(performance::check_model(model))
  print(r.squaredGLMM(model))
    # Save ggpredict output to object p
  p <- ggeffects::ggpredict(model, terms = terms)
  print(plot(p) + ggplot2::scale_color_discrete())
}
```

## Get predictor means
```{r}
get_data_means <- function(data, predictors_list, response){
  data %>%
    group_by(across(all_of(predictors_list))) %>%
    summarise(!!paste0(response, "_mean") := mean(.data[[response]], na.rm = TRUE),
              .groups = "drop")
}
```

## Make a function to run the the models for likelihood ratio tests where REML must be FALSE
```{r}
compare_lmer_models <- function(data, response, predictors_list, random_effects) {
  # --- NEW CLEANING: remove NA/NaN/Inf ONLY for the response variable ---
  data_clean <- data %>%
    filter(is.finite(.data[[response]])) %>%   # drop NA / NaN / Inf in response
    tidyr::drop_na(all_of(predictors_list)) %>%   # drop rows missing predictor values
    droplevels()
  
  # Build random-effects string
  re_str <- paste(paste0("(1 | ", random_effects, ")"), collapse = " + ")
  
  # Full model (REML = FALSE required for LRTs)
  fixed_all <- paste(predictors_list, collapse = " + ")
  f_full <- as.formula(paste(response, "~", fixed_all, "+", re_str))
  model_all <- lme4::lmer(f_full, data = data_clean, REML = FALSE)

  # Refit full model with REML = TRUE for reporting coefficients and diagnostics
  model_all_reml <- lme4::lmer(f_full, data = data_clean, REML = TRUE)

  # Drop-one models + LRTs
  models <- list()
  anova_result <- list()

  for (dropped in predictors_list) {
    fixed_reduced <- paste(setdiff(predictors_list, dropped), collapse = " + ")
    f_reduced <- as.formula(paste(response, "~", fixed_reduced, "+", re_str))
    models[[dropped]] <- lme4::lmer(f_reduced, data = data_clean, REML = FALSE)
    anova_result[[dropped]] <- anova(models[[dropped]], model_all, test = "Chisq")
  }

  data_means <- get_data_means(data_clean, predictors_list, response)

  list(
    models = models,          # named by the predictor that was dropped
    model_all = model_all,    # ML fit — use only for LRTs
    model_all_reml = model_all_reml,  # REML fit — use for reporting and diagnostics
    anova_result = anova_result,
    data_means = data_means,
    used_data = data_clean
  )
}

predictors_list <- c("salinity", "nitrate", "temp")
predictors_list_species <- c("salinity", "nitrate", "species") # for species comparisons
```
## Fv/Fm________________________________________________________________
### Inputs for Ulva Fv/Fm
```{r}
ulva_ww_fvfm_results <- compare_lmer_models(
  data = ulva_ww,
  response = "fv_fm",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results
### Full model
```{r}
summary(ulva_ww_fvfm_results$model_all_reml)
```

```{r}
# Does salinity effect Ulva FvFm?
ulva_ww_fvfm_results$anova_result[["salinity"]]
# Does nitrate effect Ulva FvFm?
ulva_ww_fvfm_results$anova_result[["nitrate"]]
#Does temp effect Ulva FvFm?
ulva_ww_fvfm_results$anova_result[["temp"]]
```

### Check model fit and get table of means
```{r}
check_model_fit(ulva_ww_fvfm_results$model_all_reml, terms = predictors_list)
ulva_ww_fvfm_results$data_means
```

### Inputs for Hypnea Fv/Fm
```{r}
hypnea_ww_fvfm_results <- compare_lmer_models(
  data = hypnea_ww,
  response = "fv_fm",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results
### Full model
```{r}
summary(hypnea_ww_fvfm_results$model_all_reml)
```

```{r}
# Does salinity effect hypnea FvFm?
hypnea_ww_fvfm_results$anova_result[["salinity"]]
# Does nitrate effect hypnea FvFm?
hypnea_ww_fvfm_results$anova_result[["nitrate"]]
#Does temp effect hypnea FvFm?
hypnea_ww_fvfm_results$anova_result[["temp"]]

```

### Check model fit
```{r}
check_model_fit(hypnea_ww_fvfm_results$model_all_reml, terms = predictors_list)
hypnea_ww_fvfm_results$data_means
```

# Pmax_____________________________________
## Inputs for ulva_ps/Pmax
```{r}
ulva_ww_pmax_results <- compare_lmer_models(
  data = ulva_ww,
  response = "pmax",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Ulva Pmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_pmax_results$model_all_reml)

# Drop-one summaries to inspect each
summary(ulva_ww_pmax_results$models[["salinity"]])
summary(ulva_ww_pmax_results$models[["nitrate"]])
summary(ulva_ww_pmax_results$models[["temp"]])

# Does salinity effect Ulva Pmax?
ulva_ww_pmax_results$anova_result[["salinity"]]
# Does nitrate effect Ulva Pmax?
ulva_ww_pmax_results$anova_result[["nitrate"]]
#Does temp effect Ulva Pmax?
ulva_ww_pmax_results$anova_result[["temp"]]

check_model_fit(ulva_ww_pmax_results$model_all_reml, terms = predictors_list)
 ulva_ww_pmax_results$data_means
```


## Inputs for Hypnea_ww/Pmax
```{r}
hypnea_ww_pmax_results <- compare_lmer_models(
  data = hypnea_ww,
  response = "pmax",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```
### Access results Hypnea Pmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_pmax_results$model_all_reml)
# Does salinity effect Hypnea Pmax?
hypnea_ww_pmax_results$anova_result[["salinity"]]
# Does nitrate effect Hypnea Pmax?
hypnea_ww_pmax_results$anova_result[["nitrate"]]
#Does temp effect Hypnea Pmax?
hypnea_ww_pmax_results$anova_result[["temp"]]

check_model_fit(hypnea_ww_pmax_results$model_all_reml, terms = predictors_list)
hypnea_ww_pmax_results$data_means
```

## Inputs for Pmax (compare both species)
```{r}
ww_pmax_results <- compare_lmer_models(
  data = ww_lmm_d9,
  response = "pmax",
  predictors_list_species,
  random_effects = c("plant_id", "rlc_order1")
)
```

### Access results Species Pmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_pmax_results$model_all_reml)

# Drop-one summaries (if you want to inspect them)
summary(ww_pmax_results$models[["species"]])

# Does species effect Pmax?
ww_pmax_results$anova_result[["species"]]

check_model_fit(ww_pmax_results$model_all_reml, terms = predictors_list_species)
ww_pmax_results$data_means
```

# NPQmax______________________________
## Inputs for Ulva_ww/NPQmax
```{r}
ulva_ww_npqmax_results <- compare_lmer_models(
  data = ulva_ww,
  response = "npq_max",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Ulva NPQmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_npqmax_results$model_all_reml)
# Drop-one summaries (to inspect them)
summary(ulva_ww_npqmax_results$models[["salinity"]])
summary(ulva_ww_npqmax_results$models[["nitrate"]])
summary(ulva_ww_npqmax_results$models[["temp"]])

# Does salinity effect Ulva NPQmax?
ulva_ww_npqmax_results$anova_result[["salinity"]]
# Does nitrate effect Ulva NPQmax?
ulva_ww_npqmax_results$anova_result[["nitrate"]]
# Does temperature effect Ulva NPQmax?
ulva_ww_npqmax_results$anova_result[["temp"]]

check_model_fit(ulva_ww_npqmax_results$model_all_reml, terms = predictors_list)
ulva_ww_npqmax_results$data_means
```


## Inputs for hypnea_ww/NPQmax
```{r}
hypnea_ww_npqmax_results <- compare_lmer_models(
  data = hypnea_ww,
  response = "npq_max",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Hypnea NPQmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_npqmax_results$model_all_reml)
# Does salinity effect Hypnea NPQmax?
hypnea_ww_npqmax_results$anova_result[["salinity"]]
# Does nitrate effect Hypnea NPQmax?
hypnea_ww_npqmax_results$anova_result[["nitrate"]]
#Does temp effect Hypnea NPQmax?
hypnea_ww_npqmax_results$anova_result[["temp"]]

check_model_fit(hypnea_ww_npqmax_results$model_all_reml, terms = predictors_list)
hypnea_ww_npqmax_results$data_means
```

## Inputs for NPQmax (compare both species)
```{r}
ww_npq_max_results <- compare_lmer_models(
  data = ww_lmm_d9,
  response = "npq_max",
  predictors_list_species,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Species NPQmax
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_npq_max_results$model_all_reml)

# Drop-one summaries
summary(ww_npq_max_results$models[["species"]])

# Does species effect NPQmax?
ww_npq_max_results$anova_result[["species"]]

check_model_fit(ww_npq_max_results$model_all_reml, terms = predictors_list_species)
ww_npq_max_results$data_means
```

#deltaNPQ____________________________
## Inputs for ulva_ww/deltaNPQ
```{r}
ulva_ww_delta_npq_results <- compare_lmer_models(
  data = ulva_ww,
  response = "delta_npq",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```
### Access results Ulva deltaNPQ
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_delta_npq_results$model_all_reml)
summary(ulva_ww_delta_npq_results$models[["nitrate"]])
summary(ulva_ww_delta_npq_results$models[["salinity"]]) 
summary(ulva_ww_delta_npq_results$models[["temp"]])

# Does salinity effect Ulva deltaNPQ?
ulva_ww_delta_npq_results$anova_result[["salinity"]]
# Does nitrate effect Ulva deltaNPQ?
ulva_ww_delta_npq_results$anova_result[["nitrate"]]
# Does temperature effect Ulva deltaNPQ?
ulva_ww_delta_npq_results$anova_result[["temp"]]

check_model_fit(ulva_ww_delta_npq_results$model_all_reml, terms = predictors_list)
ulva_ww_delta_npq_results$data_means
```

## Inputs for hypnea_ww/deltaNPQ
```{r}
hypnea_ww_delta_npq_results <- compare_lmer_models(
  data = hypnea_ww,
  response = "delta_npq",
  predictors_list,
  random_effects = c("run_combo") # plant_id and rlc_order have zero effect
)
```

### Access results Hypnea deltaNPQ
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_delta_npq_results$model_all_reml)
summary(hypnea_ww_delta_npq_results$models[["nitrate"]])
summary(hypnea_ww_delta_npq_results$models[["salinity"]]) 
summary(hypnea_ww_delta_npq_results$models[["temp"]])

# Does salinity effect Hypnea deltaNPQ?
hypnea_ww_delta_npq_results$anova_result[["salinity"]]
# Does nitrate effect Hypnea deltaNPQ?
hypnea_ww_delta_npq_results$anova_result[["nitrate"]]
# Does temperature effect Hypnea deltaNPQ?
hypnea_ww_delta_npq_results$anova_result[["temp"]]

check_model_fit(hypnea_ww_delta_npq_results$model_all_reml, terms = predictors_list)
hypnea_ww_delta_npq_results$data_means
```

#Ek____________________________________________
## Inputs for ulva_ww/Ek
```{r}
ulva_ww_ek_results <- compare_lmer_models(
  data = ulva_ww,
  response = "ek_est",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Ulva Ek
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ulva_ww_ek_results$model_all_reml)
summary(ulva_ww_ek_results$models[["nitrate"]])
summary(ulva_ww_ek_results$models[["salinity"]])
summary(ulva_ww_ek_results$models[["temp"]])
# Does salinity effect Ulva Ek?
ulva_ww_ek_results$anova_result[["salinity"]]
# Does nitrate effect Ulva Ek?
ulva_ww_ek_results$anova_result[["nitrate"]]
# Does temp effect Ulva Ek?
ulva_ww_ek_results$anova_result[["temp"]]

check_model_fit(ulva_ww_ek_results$model_all_reml, terms = predictors_list)
ulva_ww_ek_results$data_means
```

## Inputs for hypnea_ww/Ek
```{r}
hypnea_ww_ek_results <- compare_lmer_models(
  data = hypnea_ww,
  response = "ek_est",
  predictors_list,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Hypnea Ek
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(hypnea_ww_ek_results$model_all_reml)
summary(hypnea_ww_ek_results$models[["nitrate"]])
summary(hypnea_ww_ek_results$models[["salinity"]])
summary(hypnea_ww_ek_results$models[["temp"]])
# Does salinity effect Hypnea Ek?
hypnea_ww_ek_results$anova_result[["salinity"]]
# Does nitrate effect Hypnea Ek?
hypnea_ww_ek_results$anova_result[["nitrate"]]
# Does temp effect Hypnea Ek?
hypnea_ww_ek_results$anova_result[["temp"]]

check_model_fit(hypnea_ww_ek_results$model_all_reml, terms = predictors_list)
hypnea_ww_ek_results$data_means
```

## Inputs for Ek (compare both species)
```{r}
ww_ek_results <- compare_lmer_models(
  data = ww_lmm_d9,
  response = "ek_est",
  predictors_list_species,
  random_effects = c("plant_id", "run_combo", "rlc_order1")
)
```

### Access results Species Ek
```{r, fig.width = 8, fig.height = 9, results ='markup'}
summary(ww_ek_results$model_all_reml)

# Drop-one summaries (if you want to inspect them)
summary(ww_ek_results$models[["species"]])

# Does species effect Ek?
ww_ek_results$anova_result[["species"]]

check_model_fit(ww_ek_results$model_all_reml, terms = predictors_list_species)
ww_ek_results$data_means
```

# HISTOGRAMS and PLOTS_____________________________________
# Function for raw data plots
```{r}
raw_plots <- function(data, response, response2, label, pretty_color, aescolor, x, x2, y, y2, title, title2, subtitle, ylim1, ylim2, 
                      color1, color2, color3, yint, vjust_t, hjust_t, vjust_s, hjust_s, labels1) {
  histo <- data %>%
    ggplot(aes(x = {{response}})) +
    geom_histogram(fill = pretty_color, color = "black", linewidth = 0.25, alpha = 0.85) +
    labs(title = label) +
    theme_bw()
  
  plot <- data %>% 
    ggplot(aes(x = nitrate, y = {{response}})) + 
    geom_boxplot(size=0.5) + 
    geom_point(alpha = 0.75, size = 3, aes(color = {{aescolor}}), position = "jitter", show.legend = TRUE) + 
    labs(x = x, y = y, title = title, subtitle = subtitle) + 
    #scale_x_discrete(labels = c("0.5μmolN", "2μmolN", "4μmolN", "8μmolN")) + 
    ylim(ylim1, ylim2) + stat_mean() + 
    scale_color_manual(values = c(color1, color2, color3)) +
    scale_x_discrete(labels = labels1) +
    geom_hline(yintercept=yint, color = "red", linewidth = 0.5, alpha = 0.5) +
    theme_bw() +
    theme(legend.position = c(0.90,0.80), plot.title = element_text(face = "italic", vjust = vjust_t, hjust = hjust_t), 
          plot.subtitle = element_text(face = "bold", size = 14, vjust = vjust_s, hjust = hjust_s))
  
 # lin_regr <- data %>%
  #  ggplot(aes(x = {{response}}, 
   #            y = {{response2}})) + 
  #  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  #  geom_smooth(method = "lm", col = "black") + 
  #  theme_bw() + 
  #  labs(title = title2, 
   #      x = x2, 
    #     y = y2) + 
  #  stat_regline_equation(label.x = 1, label.y = max(data$response2) * 1.1) + 
  #  stat_cor(label.x = 1, label.y = max(data$response2) * 1.05)
  
  return(list(
    histo = histo,
    plot = plot)) 
}
```

# INPUTS for histograms and boxplots-----------------
### FvFm

### Inputs for Ulva/fvfm
```{r}
fvfm_ulva <- raw_plots(
  data = ulva_ww,
  response = fv_fm,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "goldenrod1",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Fv/Fm (AU)",
  y = "Day 9 Fv/Fm (AU)",
  y2 = "9-Day Growth (%)",
  title = "A - Ulva lactuca",
  title2 = "Ulva lactuca --- FvFm (AU) vs 9-Day Growth (%)",
  subtitle = "Fv/Fm",
  ylim1 = 0, 
  ylim2 = 1,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = .5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(fvfm_ulva$histo)
plot(fvfm_ulva$plot)
#ggsave("pmax_ulva_sgdww.png", path = "wastewater/plots/", 
#      width = 7, height = 6, units = "in", dpi = 300, scale = 1)
```

### Inputs for hypnea/fvfm
```{r}
fvfm_hypnea_ww <- raw_plots(
  data = hypnea_ww,
  response = fv_fm,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon4",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Fv/Fm (AU)",
  y = "Day 9 Fv/Fm (AU)",
  y2 = "9-Day Growth (%)",
  title = "B - Hypnea musciformis",
  title2 = "Hypnea musciformis --- Fv/Fm vs 9-Day Growth (%)",
  subtitle = "Fv/Fm",
  ylim1 = 0, 
  ylim2 = 1,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 0.5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(fvfm_hypnea_ww$histo)
plot(fvfm_hypnea_ww$plot)
```

#Pmax ------------------------

# Inputs for Ulva/pmax
```{r}
pmax_ulva <- raw_plots(
  data = ulva_ww,
  response = pmax,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "goldenrod1",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Pmax (μmols electrons m-2 s-1)",
  y = "Day 9 Pmax (μmols electrons m-2 s-1)",
  y2 = "9-Day Growth (%)",
  title = "A - Ulva lactuca",
  title2 = "Ulva lactuca --- Pmax vs 9-Day Growth (%)",
  subtitle = "Pmax",
  ylim1 = 0, 
  ylim2 = 135,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 20,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(pmax_ulva$histo)
plot(pmax_ulva$plot)
```

# Inputs for hypnea/pmax
```{r}
pmax_hypnea_ww <- raw_plots(
  data = hypnea_ww,
  response = pmax,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon4",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Pmax (μmols electrons m-2 s-1)",
  y = "Day 9 Pmax (μmols electrons m-2 s-1)",
  y2 = "9-Day Growth (%)",
  title = "B - Hypnea musciformis",
  title2 = "Hypnea musciformis --- Pmax vs 9-Day Growth (%)",
  subtitle = "Pmax",
  ylim1 = 0, 
  ylim2 = 135,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 20,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(pmax_hypnea_ww$histo)
plot(pmax_hypnea_ww$plot)
```

#NPQmax----------------

# Inputs for ulva_ww/NPQmax
```{r}
npqmax_ulva_ww <- raw_plots(
  data = ulva_ww,
  response = npq_max,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "darkolivegreen2",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "NPQmax",
  y = "Day 9 NPQmax",
  y2 = "9-Day Growth (%)",
  title = "C - Ulva lactuca",
  title2 = "Ulva lactuca --- NPQmax vs 9-Day Growth (%)",
  subtitle = "NPQmax",
  ylim1 = -0.5, 
  ylim2 = 2,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 0.5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(npqmax_ulva_ww$histo)
plot(npqmax_ulva_ww$plot)
#ggsave("npqmax_ulva_sgdww.png", path = "wastewater/plots/", 
 #      width = 7, height = 6, units = "in", dpi = 300, scale = 1)

#inputs for hypnea_ww/NPQmax
npqmax_hypnea_ww <- raw_plots(
  data = hypnea_ww,
  response = npq_max,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon1",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "NPQmax",
  y = "Day 9 NPQmax",
  y2 = "9-Day Growth (%)",
  title = "D - Hypnea musciformis",
  title2 = "Hypnea musciformis --- NPQmax vs 9-Day Growth (%)",
  subtitle = "NPQmax",
  ylim1 = -0.5, 
  ylim2 = 2,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 0.5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(npqmax_hypnea_ww$histo)
plot(npqmax_hypnea_ww$plot)
```

#deltaNPQ----------------

# Inputs for ulva_ww/deltaNPQ
```{r}
delta_npq_ulva_ww <- raw_plots(
  data = ulva_ww,
  response = delta_npq,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "darkolivegreen2",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "ΔNPQ",
  y = "Day 9 ΔNPQ",
  y2 = "9-Day Growth (%)",
  title = "E - Ulva lactuca",
  title2 = "Ulva lactuca --- ΔNPQ vs 9-Day Growth (%)",
  subtitle = "ΔNPQ",
  ylim1 = 0, 
  ylim2 = 2,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 0.5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(delta_npq_ulva_ww$histo)
plot(delta_npq_ulva_ww$plot)
```

# Inputs for hypnea_ww/deltaNPQ
```{r}
delta_npq_hypnea_ww <- raw_plots(
  data = hypnea_ww,
  response = delta_npq,
  response2 = d9_growth_percent,
  label = "Hypnea musciformis",
  pretty_color = "maroon3",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "ΔNPQ",
  y = "Day 9 ΔNPQ",
  y2 = "9-Day Growth (%)",
  title = "F - Hypnea musciformis",
  title2 = "Hypnea musciformis --- ΔNPQ vs 9-Day Growth (%)",
  subtitle = "ΔNPQ",
  ylim1 = 0, 
  ylim2 = 2,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 0.5,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(delta_npq_hypnea_ww$histo)
plot(delta_npq_hypnea_ww$plot)
```


#Ek----------------

# Inputs for ulva_ww/Ek
```{r}
ek_ulva_ww <- raw_plots(
  data = ulva_ww,
  response = ek_est,
  response2 = growth,
  label = "Ulva lactuca",
  pretty_color = "goldenrod4",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Ek (μmols photons m-2 s-1)",
  y = "Day 9 Ek (μmols photons m-2 s-1)",
  y2 = "9-Day Growth (%)",
  title = "G - Ulva lactuca",
  title2 = "Ulva lactuca --- Ek vs 9-Day Growth (%)",
  subtitle = "Ek",
  ylim1 = 0, 
  ylim2 = 225,
  color1 = "forestgreen",
  color2 = "darkolivegreen3",
  color3 = "olivedrab4",
  yint = 50,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(ek_ulva_ww$histo)
plot(ek_ulva_ww$plot)
#ggsave("ek_ulva_sgdww.png", path = "wastewater/plots/", 
 #      width = 7, height = 6, units = "in", dpi = 300, scale = 1)
```

## Inputs for hypnea_ww/Ek
```{r}
ek_hypnea_ww <- raw_plots(
  data = hypnea_ww,
  response = ek_est,
  response2 = growth,
  label = "Hypnea musciformis",
  pretty_color = "maroon3",
  aescolor = salinity,
  x = "Nitrate",
  x2 = "Ek (μmols photons m-2 s-1)",
  y = "Day 9 Ek (μmols photons m-2 s-1)",
  y2 = "9-Day Growth (%)",
  title = "H - Hypnea musciformis",
  title2 = "Hypnea musciformis --- Ek vs 9-Day Growth (%)",
  subtitle = "Ek",
  ylim1 = 0, 
  ylim2 = 225,
  color1 = "deeppink4",
  color2 = "deeppink3",
  color3 = "hotpink",
  yint = 50,
  vjust_t = -15,
  hjust_t = 0.05,
  vjust_s = -10,
  hjust_s = 0.95,
  labels1 = c("80 μmol", 
              "245 μmol",
              "748 μmol",
              "2287 μmol")
)
plot(ek_hypnea_ww$histo)
plot(ek_hypnea_ww$plot)
```

### Tables________________________________________________________________

```{r}
print_model <- function(model, species, response) {
  sjPlot::tab_model(
    model,
    title = paste(species, "-", response),
    show.intercept = TRUE,
    show.se = TRUE,
    show.stat = TRUE,
    show.df = TRUE
  )
}
print_model(ulva_ww_pmax_results$model_all_reml, "Ulva", "Pmax")
print_model(hypnea_ww_pmax_results$model_all_reml, "Hypnea", "Pmax")

print_model(ulva_ww_npqmax_results$model_all_reml, "Ulva", "NPQmax")
print_model(hypnea_ww_npqmax_results$model_all_reml, "Hypnea", "NPQmax")

print_model(ulva_ww_delta_npq_results$model_all_reml, "Ulva", "delta NPQ")
print_model(hypnea_ww_delta_npq_results$model_all_reml, "Hypnea", "delta NPQ")

print_model(ulva_ww_ek_results$model_all_reml, "Ulva", "Ek")
print_model(hypnea_ww_ek_results$model_all_reml, "Hypnea", "Ek")
```

