---
title: "Acanthophora Growth and Apices Script"
author: "Angela Richards Donà"
date: "01/05/2024"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---
## Acanthophora spicifera GROWTH RATE and APICES Analysis, Script Chunks, and Plots

This is the analysis of two runs of Acanthophora spicifera salinity and nutrient experiments 
conducted on the lanai in St. John 616 in November and December 2023. These experiments incorporated two salinity/nutrient levels and two temperature levels.

Packages loaded:

```{r message=FALSE}
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
library(strengejacke)
#library(mmtable2) no longer available
library(gt)
library(purrr)
library(stringr)
library(tidyr)
library(piecewiseSEM)
library(easystats)
library(magrittr)
```


Open weight dataset and make columns for growth rate from initial and final weights
```{r}
acan_growth <- read.csv("../input/acanthophora_growth_r1.csv")
```

Make a new column for weight change (difference final from initial)
```{r}
acan_growth$growth_rate_percent <- (acan_growth$final_weight - acan_growth$initial_weight) / acan_growth$initial_weight * 100
```

Make a new column for daily growth rate from 8 day study (steady growth rate assumed rather than exponential)
```{r}
acan_growth$steady_growth_daily <- acan_growth$growth_rate_percent / 8
```

Assigns temperature as a factor
```{r}
acan_growth$temperature <- as.factor(acan_growth$temperature)
```

Assigns treatment as characters from integers then to factors
```{r}
acan_growth$treatment <- as.factor(as.character(acan_growth$treatment))
```

Assign run as a factor
```{r}
acan_growth$run <- as.factor(acan_growth$run)
```

Assign plant ID as a factor
```{r}
acan_growth$plant_ID <- as.factor(acan_growth$plant_ID)
```

Assign RLC order as a factor
```{r}
acan_growth$rlc_order <- as.factor(acan_growth$rlc_order)
```

Create subsets for the plots and remove one datapoint that is strongly skewing data (overall result does not change)
```{r}

acan_growth <- subset(acan_growth, secondary_apices != "200")

acan_growth$treatment_graph[acan_growth$treatment == 0] <- "1) 35ppt/0.5umol"
acan_growth$treatment_graph[acan_growth$treatment == 1] <- "2) 28ppt/14umol" 

acan_growth$temperature_graph[acan_growth$temperature == 24] <- "24°C"
acan_growth$temperature_graph[acan_growth$temperature == 28] <- "28°C"
```

Make a histogram of the growth rate data
```{r}
hist(acan_growth$growth_rate_percent, main = paste("Acanthophora spicifera 9-Day Growth (%)"), col = "maroon", labels = TRUE)
#or
acan_growth %>% ggplot(aes(growth_rate_percent)) +
  geom_histogram(binwidth=5, fill = "#990066", color = "black", linewidth = 0.25, alpha = 0.85) +
  theme_bw()
```

##Run model without RLC_order as this has little effect, add initial number of apices
```{r}
growth_model_acan <- lmer(formula = growth_rate_percent ~ treatment + temperature +
                            (1 | plant_ID) + (1 | run) + (1 | initial_axes), data = acan_growth, REML = TRUE)
```

Plot histograms and performance checks
```{r}
hist(resid(growth_model_acan))
plot(resid(growth_model_acan) ~ fitted(growth_model_acan))
qqnorm(resid(growth_model_acan))
qqline(resid(growth_model_acan))

performance::check_model(growth_model_acan)
```

Get R2 and summarize
```{r}
rsquared(growth_model_acan)
summary(growth_model_acan)
```

View random effects levels and make a table of the model output
```{r}
ranef(growth_model_acan)
tab_model(growth_model_acan, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```

Plot the model effects
```{r}
plot(allEffects(growth_model_acan))
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
acan_growth_treatment_null <- lmer(formula = growth_rate_percent ~ temperature + (1 | plant_ID) + (1 | run), data = acan_growth, REML = FALSE)
acan_growth_model2 <- lmer(formula = growth_rate_percent ~ treatment + temperature + (1 | plant_ID) + (1 | run), data = acan_growth, REML = FALSE)
anova(acan_growth_treatment_null, acan_growth_model2)
acan_growth_temperature_null <- lmer(formula = growth_rate_percent ~ treatment + (1 | plant_ID) + (1 | run), data = acan_growth, REML = FALSE)
acan_growth_model3 <- lmer(formula = growth_rate_percent ~ treatment + temperature + (1 | plant_ID) + (1 | run), data = acan_growth, REML = FALSE)
anova(acan_growth_temperature_null, acan_growth_model3)
```

Simple boxplots for treatment and temperature
```{r}
acan_growth %>% ggplot(aes(treatment_graph, growth_rate_percent)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = temperature), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "9-Day Growth (%)", title= "A", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("T0) 35 ppt/0.5 μmol N", "T1) 28 ppt/14 μmol N")) + 
  ylim(-60, 60) + stat_mean() + 
  scale_color_manual(values = c("azure4", "darkgoldenrod1")) +
  geom_hline(yintercept=0, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```
```{r}
acan_growth %>% ggplot(aes(temperature_graph, growth_rate_percent)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = treatment), position = "jitter", show.legend = TRUE) + 
  labs(x="Temperature", y= "9-day Growth (%)", title= "B", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("24°C", "28°C")) + 
  ylim(-60, 50) + stat_mean() + 
  geom_hline(yintercept=0, color = "red", linewidth = 0.5, alpha = 0.5) +
  scale_color_manual(values = c("azure3", "darkgoldenrod")) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means
```{r}
acan_growth %>% group_by(treatment) %>% summarise_at(vars(growth_rate_percent), list(mean = mean))
acan_growth %>% group_by(temperature) %>% summarise_at(vars(growth_rate_percent), list(mean = mean))
```



##RUN MODEL FOR SECONDARY APICES

First plot histogram
```{r}
acan_growth %>% ggplot(aes(secondary_apices)) +
  geom_histogram(binwidth=5, fill = "#990066", color = "black", linewidth = 0.25, alpha = 0.85) +
  theme_bw()
```

Run model for secondary apices
```{r}
apices_model_acan <- lmer(formula = secondary_apices ~ treatment + temperature +
                            (1 | plant_ID) + (1 | run) + (1 | rlc_order), data = acan_growth, REML = TRUE)
```

Check performance of model
```{r}
hist(resid(apices_model_acan))
plot(resid(apices_model_acan) ~ fitted(apices_model_acan))
qqnorm(resid(apices_model_acan))
qqline(resid(apices_model_acan))

performance::check_model(apices_model_acan)
```

R squared
```{r}
rsquared(apices_model_acan)
summary(apices_model_acan)
```

View random effects levels, print table and plot all effects
```{r}
ranef(apices_model_acan)
tab_model(apices_model_acan, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(apices_model_acan))
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
acan_apices_treatment_null <- lmer(formula = secondary_apices~ temperature + (1 | plant_ID) + (1 | run) + (1 | rlc_order), data = acan_growth, REML = FALSE)
acan_apices_model2 <- lmer(formula = secondary_apices ~ treatment + temperature + (1 | plant_ID) + (1 | run) + (1 | rlc_order), data = acan_growth, REML = FALSE)
anova(acan_apices_treatment_null, acan_apices_model2)
```
```{r}
acan_apices_temperature_null <- lmer(formula = secondary_apices ~ treatment + (1 | plant_ID) + (1 | run) + (1 | rlc_order), data = acan_growth, REML = FALSE)
acan_apices_model3 <- lmer(formula = secondary_apices ~ treatment + temperature + (1 | plant_ID) + (1 | run) + (1 | rlc_order), data = acan_growth, REML = FALSE)
anova(acan_apices_temperature_null, acan_apices_model3)
```

Plots
```{r}
acan_growth %>% ggplot(aes(treatment_graph, secondary_apices)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = temperature), position = "jitter", show.legend = TRUE) + 
  labs(x="treatment", y= "Number of Secondary Apices", title= "A", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("T0) 35 ppt/0.5 μmol N", "T1) 28 ppt/14 μmol N")) + 
  ylim(-1, 130) + stat_mean() + 
  scale_color_manual(values = c("azure4", "darkgoldenrod1")) +
  geom_hline(yintercept=0, color = "red", linewidth = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))


acan_growth %>% ggplot(aes(temperature_graph, secondary_apices)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = treatment), position = "jitter", show.legend = TRUE) + 
  labs(x="Temperature", y= "Number of Secondary Apices", title= "B", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("24°C", "28°C")) + 
  ylim(-1, 130) + stat_mean() + 
  geom_hline(yintercept=0, color = "red", linewidth = 0.5, alpha = 0.5) +
  scale_color_manual(values = c("azure3", "darkgoldenrod")) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means
```{r}
acan_growth %>% group_by(treatment) %>% summarise_at(vars(secondary_apices), list(mean = mean))
acan_growth %>% group_by(temperature) %>% summarise_at(vars(secondary_apices), list(mean = mean))
```






