---
title: "Acanthophora spicifera Hsat model"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output:
  html_document: 
    fig_width: 9
    fig_height: 9
---
This script runs a mixed effects model analysis of Hsat in minutes to see if treatment or temperature fixed effects had influence on this dependent variable.  
Supersat is the time above Ek for each individual in minutes, these are binned into three time periods from day 1 to day 9
Supersat avg is the mean of the three time periods

# Load the various libraries
```{r, message = FALSE}
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
#library(mmtable2)
library(gt)
library(purrr)
library(stringr)
library(tidyr)
library(magrittr)
```

Read in the data
```{r}
ek_irrad_data <- read.csv("../transformed/acan_hsat_dspi.csv")
```

Assign variables as factors, characters, numeric, as needed
```{r}
ek_irrad_data$run <- as.factor(ek_irrad_data$run)
ek_irrad_data$temperature <- as.factor(as.character(ek_irrad_data$temperature))
ek_irrad_data$treatment <- as.factor(as.character(ek_irrad_data$treatment))
```

For plots-match up number of treatment with detailed
```{r}
ek_irrad_data$treatment_graph[ek_irrad_data$treatment == 0] <- "1) 35ppt/0.5umol"
ek_irrad_data$treatment_graph[ek_irrad_data$treatment == 1] <- "2) 28ppt/14umol" 
```

For plots - same but for temp
```{r}
ek_irrad_data$temperature_graph[ek_irrad_data$temperature == 24] <- "1) 24 °C"
ek_irrad_data$temperature_graph[ek_irrad_data$temperature == 28] <- "2) 28 °C" 
```

Make a histogram of the data 
```{r}
hist(ek_irrad_data$supersat_avg, main = paste("Acanthophora spicifera"), col = "maroon", labels = TRUE)

ek_irrad_data %>% ggplot(aes(supersat_avg)) +
        geom_histogram(binwidth=25, fill = "maroon", color = "black", linewidth = 0.25, alpha = 0.85) +
        theme_bw()
```

Run model without interaction between the treatments and temperature - supersat_avg is in minutes
```{r}
hsat_model <- lmer(formula = supersat_avg ~ treatment + temperature + (1 | run) + (1 | plantID) + (1 | rlc_order), data = ek_irrad_data)
```
Check residual plots
```{r}
hist(resid(hsat_model))
plot(resid(hsat_model) ~ fitted(hsat_model))
qqnorm(resid(hsat_model))
qqline(resid(hsat_model))
```
Check the performance of the model
```{r}
performance::check_model(hsat_model)
r.squaredGLMM(hsat_model)
summary(hsat_model)
plot(allEffects(hsat_model))
tab_model(hsat_model, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```
Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
hsat_treatment_null <- lmer(formula = supersat_avg ~ temperature + (1 | run) + (1 |plantID) + (1 | rlc_order), data = ek_irrad_data, REML = FALSE)
hsat_model2 <- lmer(formula = supersat_avg ~ treatment + temperature + (1 | run) + (1 |plantID) + (1 | rlc_order), data = ek_irrad_data, REML = FALSE)
anova(hsat_treatment_null, hsat_model2)
hsat_temperature_null <- lmer(formula = supersat_avg ~ treatment + (1 | run) + (1 |plantID) + (1 | rlc_order), data = ek_irrad_data, REML = FALSE)
hsat_model3 <- lmer(formula = supersat_avg ~ treatment + temperature + (1 | run) + (1 |plantID) + (1 | rlc_order), data = ek_irrad_data, REML = FALSE)
anova(hsat_temperature_null, hsat_model3)
```

Plot the outcomes
```{r}
ek_irrad_data %>% ggplot(aes(treatment_graph, supersat_avg)) + 
        geom_boxplot(size=0.5) + 
        geom_point(alpha = 0.75, size = 3, aes(color = temperature), position = "jitter", show.legend = TRUE) + 
        labs(x="Treatment (salinty/nitrate)", y= "Hsat Time (min)", title= "A", subtitle = "Acanthophora spicifera") + 
        scale_x_discrete(labels = c("35ppt/0.5umolN", "28ppt/14umolN")) + 
        ylim(0, 650) + stat_mean() + 
        geom_hline(yintercept=200, color = "red", linewidth = 0.5, alpha = 0.5) +
        scale_color_manual(values = c("#9C0627", "#BB589F")) +
        theme_bw() +
        theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))


ek_irrad_data %>% ggplot(aes(temperature_graph, supersat_avg)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 3, aes(color = treatment), position = "jitter", show.legend = TRUE) + 
  labs(x="Temperature (°C)", y= "Hsat Time (min)", title= "B", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("24 °C", "28 °C")) + 
  ylim(0, 650) + stat_mean() + 
  geom_hline(yintercept=200, color = "red", linewidth = 0.5, alpha = 0.5) +
  scale_color_manual(values = c("#BB589F", "#F4B4E2")) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for variables of interest
```{r}
ek_irrad_data %>% group_by(treatment) %>% summarise_at(vars(supersat_avg), list(mean = mean))
ek_irrad_data %>% group_by(temperature) %>% summarise_at(vars(supersat_avg), list(mean = mean))
ek_irrad_data %>% group_by(run) %>% summarise_at(vars(day_length_avg), list(mean = mean))
```

Add growth rate from other dataset to this one and subset by species for regression
```{r}
growth_rate <- read.csv("../input/acanthophora_growth_r1.csv")
growth_rate$treatment <- as.factor(growth_rate$treatment)
growth_rate$growth_rate_percent <- (growth_rate$final_weight - growth_rate$initial_weight) / growth_rate$initial_weight * 100
ek_irrad_data$growth_rate_percent <- growth_rate$growth_rate_percent
ek_irrad_data$apices <- growth_rate$mean_apices_per_axis
```

Plot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
acan_apices_hsat_graph <- ggplot(ek_irrad_data, aes(x=supersat_avg, y=apices)) + 
        geom_point(alpha = 0.5, size = 2, show.legend = TRUE, aes(color = treatment)) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        labs(title = "Acanthophora spicifera Hsat vs Mean # Apices per Axis", x = "Hsat time (minutes)", 
             y = "mean apices/axis") + stat_regline_equation(label.x = 100, label.y = 30) + stat_cor()
acan_apices_hsat_graph
```
```{r}
acan_growth_hsat_graph <- ggplot(ek_irrad_data, aes(x=supersat_avg, y=growth_rate_percent)) + 
        geom_point(alpha = 0.5, size = 2, show.legend = TRUE, aes(color = treatment)) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        labs(title = "Acanthophora spicifera Hsat vs Growth Rate", x = "Hsat time (minutes)", 
             y = "growth rate (%)") + stat_regline_equation(label.x = 100, label.y = 30) + stat_cor()
acan_growth_hsat_graph
```