---
title: "Acanthophora Pmax & NPQ Model Script"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output: 
html_document: 
    fig_width: 9
    fig_height: 9
---
## Acanthophora spicifera PHOTOSYNTHESIS (Pmax and NPQ) Analysis, Script Chunks, and Plots

This is the analysis of two runs of Acanthophora spicifera salinity and nutrient experiments 
conducted on the lanai in St. John 616 in November and December 2023. These experiments incorporated two salinity/nutrient levels and two temperature levels. 
Pmax, maximum NPQ and delta NPQ are used for these analyses. The max NPQ used is that which also satisfies the need for yield to be greater than 0.1. deltaNPQ takes that maximum value and subtracts the initial value of NPQ that is recorded at the second sat pulse when Epar is at 66. 

Load the appropriate packages
```{r message=FALSE}
library(lme4)
library(lmerTest)
library(afex)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
#library(mmtable2)
library(gt)
library(purrr)
library(stringr)
library(tidyr)
```

Load this file for normalized to quantum efficiency of photosynthesis per Silsbe and Kromkamp
```{r}
acan_ps <- read.csv("../transformed/acan_ek_alpha_normalized_2024.csv")
```

Assign run as a factor
```{r}
acan_ps$run <- as.factor(acan_ps$run)
```

Assign temperature as a factor
```{r}
acan_ps$temp <- as.factor(acan_ps$temp)
```

Assigns treatment as characters from integers then to factors
```{r}
acan_ps$treatment <- as.factor(as.character(acan_ps$treatment))
```

Assign deltaNPQ as numerical
```{r}
acan_ps$deltaNPQ <- as.numeric(acan_ps$deltaNPQ)
```

Use Day 9 for final analysis
```{r}
acan_day9 <- subset(acan_ps, rlc_day == 9)
acan_day9$treatment_graph[acan_day9$treatment == 0] <- "1) 35ppt/0.5umol"
acan_day9$treatment_graph[acan_day9$treatment == 1] <- "2) 28ppt/14umol" 
```

```{r}
acan_day9$temperature_graph[acan_day9$temp == 24] <- "24°C"
acan_day9$temperature_graph[acan_day9$temp == 28] <- "28°C"
```
##Pmax
Plot histogram for Pmax
```{r}
acan_day9 %>% ggplot(aes(pmax)) +
  geom_histogram(binwidth=5, fill = "goldenrod1", color = "black", linewidth = 0.25, alpha = 0.85) +
  theme_bw()
```

Run model
```{r}
acan_day9_pmax_model <- lmer(formula = pmax ~ treatment + temp + (1 | run) + (1 | plantID) + (1 | rlc_order), data = acan_day9)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
acan_pmax_treatment_null <- lmer(formula = pmax ~ temp + (1 | run) + (1 | plantID) + (1 | rlc_order), data = acan_day9, REML = FALSE)
acan_pmax_model2 <- lmer(formula = pmax ~ treatment + temp + (1 | run) + (1 | plantID) + (1 | rlc_order), data = acan_day9, REML = FALSE)
anova(acan_pmax_treatment_null, acan_pmax_model2)
acan_pmax_temperature_null <- lmer(formula = pmax ~ treatment + (1 | run) + (1 | plantID) + (1 | rlc_order), data = acan_day9, REML = FALSE)
acan_pmax_model3 <- lmer(formula = pmax ~ treatment + temp + (1 | run) + (1 | plantID) + (1 | rlc_order), data = acan_day9, REML = FALSE)
anova(acan_pmax_temperature_null, acan_pmax_model3)
```

Make residual plots of the data for Acanthophora
```{r}
hist(resid(acan_day9_pmax_model))
plot(resid(acan_day9_pmax_model) ~ fitted(acan_day9_pmax_model))
qqnorm(resid(acan_day9_pmax_model))
qqline(resid(acan_day9_pmax_model))
```

Check the performance of the model
```{r fig.height=9, fig.width=9}
performance::check_model(acan_day9_pmax_model)
```
Get R squared
```{r}
r.squaredGLMM(acan_day9_pmax_model)
```

Make plots and tables for the data
```{r}
tab_model(acan_day9_pmax_model, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(acan_day9_pmax_model))
```

```{r}
acan_day9 %>% ggplot(aes(treatment_graph, pmax)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = temp), position = "jitter", show.legend = TRUE) + 
  labs(x="salinity/nitrate", y= "Day 9 Pmax (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("35ppt/0.5umolN", "28ppt/14umolN")) + 
  ylim(-1, 90) + stat_mean() + 
  scale_color_manual(values = c("brown", "brown2")) +
  geom_hline(yintercept=50, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```
```{r}
acan_day9 %>% ggplot(aes(temperature_graph, pmax)) +
  geom_boxplot(size=0.5) +
  geom_point(alpha = 0.75, size = 5, aes(color = treatment), position = "jitter", show.legend = TRUE) +
  labs(x="temperature", y= "Day 9 Pmax (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") +
  ylim(-1, 90) + stat_mean() + 
  scale_color_manual(values = c("brown4", "brown1")) +
  geom_hline(yintercept=50, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for pmax
```{r}
acan_day9 %>% group_by(treatment) %>% summarise_at(vars(pmax), list(mean = mean))
acan_day9 %>% group_by(temp) %>% summarise_at(vars(pmax), list(mean = mean))
```

##Linear regression Pmax vs Growth--------------------------------------------------------------------------------

Add growth rate from other dataset to this one and subset by species
```{r}
growth_rate <- read.csv("../input/acanthophora_growth_r1.csv")
growth_rate$treatment <- as.factor(growth_rate$treatment)
```

Make a new column for weight change (difference final from initial)
```{r}
growth_rate$growth_rate_percent <- (growth_rate$final_weight - growth_rate$initial_weight) / growth_rate$initial_weight * 100
acan_day9$growth_rate_percent <- growth_rate$growth_rate_percent
acan_day9$secondary_apices <- growth_rate$secondary_apices
acan_day9_sub <- subset(acan_day9, secondary_apices != "200")
```

Plot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
acan_growth_pmax_graph <- ggplot(acan_day9_sub, aes(x=pmax, y=growth_rate_percent)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera Pmax vs 9-Day Growth (%)", x = "Pmax (μmols electrons m-2 s-1)", 
       y = "9-Day Growth (%)") + stat_regline_equation(label.x = 10, label.y = 65) + stat_cor(label.x = 10, label.y = 60)
acan_growth_pmax_graph
```

Plot a regression between the photosynthetic independent variables of interest and apices/axis
```{r}
acan_apices_pmax_graph <- ggplot(acan_day9_sub, aes(x=pmax, y=secondary_apices)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera Pmax vs # Secondary Apices", x = "Pmax (μmols electrons m-2 s-1)", 
       y = "Secondary Apices (#)") + stat_regline_equation(label.x = 10, label.y = 65) + stat_cor(label.x = 10, label.y = 60)
acan_apices_pmax_graph
```


## NPQ analysis using max values that also satisfy the yield>0.1 argument

Plot data histogram
```{r}
acan_day9 %>% ggplot(aes(maxNPQ_Ypoint1)) +
  geom_histogram(binwidth=0.05, fill = "violet", color = "black", size = 0.25, alpha = 0.85) +
  theme_bw()
```

Run model without interaction between the treatments and temperature
```{r}
acan_day9_maxNPQ_model <- lmer(formula = maxNPQ_Ypoint1 ~ treatment + temp + (1 | run) + (1 | plantID), data = acan_day9)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
acan_maxNPQ_treatment_null <- lmer(formula = maxNPQ_Ypoint1 ~ temp + (1 | run) + (1 | plantID), data = acan_day9, REML = FALSE)
acan_maxNPQ_model2 <- lmer(formula = maxNPQ_Ypoint1 ~ treatment + temp + (1 | run) + (1 | plantID), data = acan_day9, REML = FALSE)
anova(acan_maxNPQ_treatment_null, acan_maxNPQ_model2)
acan_maxNPQ_temperature_null <- lmer(formula = maxNPQ_Ypoint1 ~ treatment + (1 | run) + (1 | plantID), data = acan_day9, REML = FALSE)
acan_maxNPQ_model3 <- lmer(formula = maxNPQ_Ypoint1 ~ treatment + temp + (1 | run) + (1 | plantID), data = acan_day9, REML = FALSE)
anova(acan_maxNPQ_temperature_null, acan_maxNPQ_model3)
```

Make residual plots of the data for Acanthophora
```{r}
hist(resid(acan_day9_maxNPQ_model))
plot(resid(acan_day9_maxNPQ_model) ~ fitted(acan_day9_maxNPQ_model))
qqnorm(resid(acan_day9_maxNPQ_model))
qqline(resid(acan_day9_maxNPQ_model))
```

Check the performance of the model
```{r fig.height=9, fig.width=9}
performance ::check_model(acan_day9_maxNPQ_model)
```

Get R-squared
```{r}
r.squaredGLMM(acan_day9_maxNPQ_model)
```

Make plots and tables for the data
```{r}
tab_model(acan_day9_maxNPQ_model, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(acan_day9_maxNPQ_model))
```

```{r}
acan_day9 %>% ggplot(aes(treatment_graph, maxNPQ_Ypoint1)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = temp), position = "jitter", show.legend = TRUE) + 
  labs(x="salinity/nitrate", y= "Day 9 NPQmax (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("35ppt/0.5umolN", "28ppt/14umolN")) + 
  ylim(0, 1) + stat_mean() + 
  scale_color_manual(values = c("brown", "brown2")) +
  geom_hline(yintercept=0.5, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))

acan_day9 %>% ggplot(aes(temperature_graph, maxNPQ_Ypoint1)) +
  geom_boxplot(size=0.5) +
  geom_point(alpha = 0.75, size = 5, aes(color = treatment), position = "jitter", show.legend = TRUE) +
  labs(x="temperature", y= "Day 9 NPQmax (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") +
  ylim(0, 1) + stat_mean() + 
  scale_color_manual(values = c("olivedrab3", "maroon4")) +
  geom_hline(yintercept=0.5, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for NPQ
```{r}
acan_day9 %>% group_by(treatment) %>% summarise_at(vars(maxNPQ_Ypoint1), list(mean = mean))
acan_day9 %>% group_by(temp) %>% summarise_at(vars(maxNPQ_Ypoint1), list(mean = mean))
```

lot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
acan_growth_maxNPQ_graph <- ggplot(acan_day9_sub, aes(x=maxNPQ_Ypoint1, y=growth_rate_percent)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera NPQmax vs 9-Day Growth (%)", x = "NPQmax (rel. units)", 
       y = "9-Day Growth (%)") + stat_regline_equation(label.x = 0.25, label.y = 38) + stat_cor(label.x = 0.25, label.y = 40)
acan_growth_maxNPQ_graph
```

Plot a regression between the photosynthetic independent variables of interest and apices/axis
```{r}
acan_apices_NPQmax_graph <- ggplot(acan_day9_sub, aes(x=maxNPQ_Ypoint1, y=secondary_apices)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera NPQmax vs # Secondary Apices", x = "NPQmax (rel. units)", 
       y = "Secondary Apices (#)") + stat_regline_equation(label.x = 0.25, label.y = 125) + stat_cor(label.x = 0.25, label.y = 128)
acan_apices_NPQmax_graph
```

## Delta NPQ

Plot histogram
```{r}
acan_day9 %>% ggplot(aes(deltaNPQ)) +
  geom_histogram(binwidth=.05, fill = "violet", color = "black", size = 0.25, alpha = 0.85) +
  theme_bw()
```

Run model without interaction between the treatments and temperature
```{r}
acan_day9_deltaNPQ_model <- lmer(formula = deltaNPQ ~ treatment + temp + (1 | plantID), data = acan_day9)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
acan_deltaNPQ_treatment_null <- lmer(formula = deltaNPQ ~ temp + (1 | plantID), data = acan_day9, REML = FALSE)
acan_deltaNPQ_model2 <- lmer(formula = deltaNPQ ~ treatment + temp + (1 | plantID), data = acan_day9, REML = FALSE)
anova(acan_deltaNPQ_treatment_null, acan_deltaNPQ_model2)
acan_deltaNPQ_temperature_null <- lmer(formula = deltaNPQ ~ treatment + (1 | plantID), data = acan_day9, REML = FALSE)
acan_deltaNPQ_model3 <- lmer(formula = deltaNPQ ~ treatment + temp + (1 | plantID), data = acan_day9, REML = FALSE)
anova(acan_deltaNPQ_temperature_null, acan_deltaNPQ_model3)
```

Make residual plots of the data for Acanthophora
```{r}
hist(resid(acan_day9_deltaNPQ_model))
plot(resid(acan_day9_deltaNPQ_model) ~ fitted(acan_day9_deltaNPQ_model))
qqnorm(resid(acan_day9_deltaNPQ_model))
qqline(resid(acan_day9_deltaNPQ_model))
```

Check the performance of the model
```{r fig.height=9, fig.width=9}
performance ::check_model(acan_day9_deltaNPQ_model)
```

Get R-squared
```{r}
r.squaredGLMM(acan_day9_deltaNPQ_model)
```

Make plots and tables for the data
```{r}
tab_model(acan_day9_deltaNPQ_model, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
plot(allEffects(acan_day9_deltaNPQ_model))
```

```{r}
acan_day9 %>% ggplot(aes(treatment_graph, deltaNPQ)) + 
  geom_boxplot(size=0.5) + 
  geom_point(alpha = 0.75, size = 5, aes(color = temp), position = "jitter", show.legend = TRUE) + 
  labs(x="salinity/nitrate", y= "Day 9 Delta NPQ (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") + 
  scale_x_discrete(labels = c("35ppt/0.5umolN", "28ppt/14umolN")) + 
  ylim(0, 0.75) + stat_mean() + 
  scale_color_manual(values = c("olivedrab2", "maroon2")) +
  geom_hline(yintercept=0.25, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))

acan_day9 %>% ggplot(aes(temperature_graph, deltaNPQ)) +
  geom_boxplot(size=0.5) +
  geom_point(alpha = 0.75, size = 5, aes(color = treatment), position = "jitter", show.legend = TRUE) +
  labs(x="temperature", y= "Day 9 Delta NPQ (μmols electrons m-2 s-1)", title= "A", subtitle = "Acanthophora spicifera") +
  ylim(0, 0.75) + stat_mean() + 
  scale_color_manual(values = c("olivedrab3", "maroon4")) +
  geom_hline(yintercept=0.25, color = "red", size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
        plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for deltaNPQ
```{r}
acan_day9 %>% group_by(treatment) %>% summarise_at(vars(deltaNPQ), list(mean = mean))
acan_day9 %>% group_by(temp) %>% summarise_at(vars(deltaNPQ), list(mean = mean))
```

Linear regression deltaNPQ vs Growth and Apices
```{r}
acan_growth_deltaNPQ_graph <- ggplot(acan_day9_sub, aes(x=deltaNPQ, y=growth_rate_percent)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera Delta NPQ vs 9-Day Growth (%)", x = " Delta NPQ (rel. units)", 
       y = "9-Day Growth (%)") + stat_regline_equation(label.x = 0.25, label.y = 38) + stat_cor(label.x = 0.25, label.y = 40)
acan_growth_deltaNPQ_graph
```

Plot a regression between the photosynthetic independent variables of interest and apices/axis
```{r}
acan_apices_deltaNPQ_graph <- ggplot(acan_day9_sub, aes(x=deltaNPQ, y=secondary_apices)) + 
  geom_point(alpha = 0.5, size = 3, show.legend = TRUE, aes(color = treatment)) + 
  geom_smooth(method = "lm", col = "black") + theme_bw() + 
  labs(title = "Acanthophora spicifera Delta NPQ vs # Secondary Apices", x = "Delta NPQ (rel. units)", 
       y = "Secondary Apices (#)") + stat_regline_equation(label.x = 0.25, label.y = 125) + stat_cor(label.x = 0.25, label.y = 128)
acan_apices_deltaNPQ_graph
```
